---
title: "Aggravating and Mitigating Role Adjustments"
output:
    word_document: 
        reference_docx: "template/template.docx"
date: "January 27, 2023"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(flextable)


amdata <- read_csv(here::here("data/io.csv")) %>% 
  filter(sentdate>="2001-10-01" | sentmonyr<="2017-09-01") %>% 
  mutate(crime_type = crime_type %>% 
           str_replace("Immigration", "Immig.") %>% 
           str_replace("Drug Trafficking", "Drug Traff."))
```


The below document performs a preliminary analysis of aggravating and mitigating role adjustments (departures through §3B1.1 and §3B1.2), as recorded in the U.S. Sentencing Commission data. Mitigating role adjustments are most common for drug trafficking crimes. Aggravating role adjustments are most commonly granted for sexual and white collar crimes. Women receive mitigating role adjustments at higher rates than men and receive aggravating role adjustments at lower rates than men.


```{r}
#The revised model (used by the Commission to examine demographic disparities in sentencing starting with the 2010 update to the Booker report) omits aggravating and mitigating role adjustments, along with a handful of other variables, out of concern that the effect of those adjustments is already captured by the presumptive sentence length (GDLINEHI).

flextable_my <- function(x){
  x %>% 
    flextable() %>% 
    width(width = .7) %>% 
    fontsize(size = 8, part = "header") %>% 
    fontsize(size = 8, part = "body")
}

amdata_tbl <- amdata %>% 
  filter(!is.na(mitigate) | !is.na(aggravate)) %>% 
  filter(!is.na(racesex) & !is.na(crime_type)) %>% 
  group_by(racesex_clean, crime_type) %>% 
  summarize(total = n(),
            n_agg = sum(aggravate, na.rm=T),
            n_mit = sum(mitigate, na.rm=T)) %>%
  mutate(pct_agg = round(n_agg/total, 3),
         pct_mit = round(n_mit/total, 3)) 

amdata_tbl %>% 
  select(crime_type, racesex_clean, n_mit) %>% 
  pivot_wider(names_from = "crime_type", values_from = n_mit) %>% 
  rename(" " = racesex_clean) %>% 
  flextable_my() %>% 
  set_caption("Number of Mitigating Role Adjustments, FY2002-FY2017")

amdata_tbl %>% 
  select(crime_type, racesex_clean, pct_mit) %>% 
  pivot_wider(names_from = "crime_type", values_from = pct_mit) %>% 
  rename(" " = racesex_clean) %>% 
  flextable_my() %>% 
  set_caption("Portion of Mitigating Role Adjustments, FY2002-FY2017")
```
<br>
<br>
 
 
```{r}
amdata_tbl %>% 
  select(crime_type, racesex_clean, n_agg) %>% 
  pivot_wider(names_from = "crime_type", values_from = n_agg) %>% 
  rename(" " = racesex_clean) %>% 
  flextable_my() %>% 
  set_caption("Number of Aggravating Role Adjustments, FY2002-FY2017")

amdata_tbl %>% 
  select(crime_type, racesex_clean, pct_agg) %>% 
  pivot_wider(names_from = "crime_type", values_from = pct_agg) %>% 
  rename(" " = racesex_clean) %>% 
  flextable_my() %>% 
  set_caption("Portion of Aggravating Role Adjustments, FY2002-FY2017")
  
  
```



```{r}
amdata_tbl %>% 
  ggplot() +
  geom_col(aes(x = racesex_clean, y = pct_mit, fill = racesex_clean))+
  theme_minimal()+
  labs(title = "Portion of Mitigating Role Adjustments by Demographic and Crime Type")+
  theme(#text = element_text(size = 20),
        title = element_text(size = 8),
        legend.position = "none",
        axis.text.x=element_text(angle=-90),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank())+
  facet_wrap(~crime_type)
```
<br>
 
```{r}

amdata_tbl %>% 
  ggplot() +
  geom_col(aes(x = crime_type, y = pct_mit, fill = crime_type))+
  theme_minimal()+
  labs(title = "Portion of Mitigating Role Adjustments by Crime Type and Demographic")+
  theme(#text = element_text(size = 20),
        title = element_text(size = 8),
        legend.position = "none",
        axis.text.x=element_text(angle=-90),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank())+
  scale_fill_brewer(palette = "Dark2")+
  facet_wrap(~racesex_clean, ncol = 2)

```

```{r}
amdata_tbl %>% 
  ggplot() +
  geom_col(aes(x = racesex_clean, y = pct_agg, fill = racesex_clean))+
  theme_minimal()+
  labs(title = "Portion of Aggravating Role Adjustments by Demographic and Crime Type")+
  theme(#text = element_text(size = 20),
        title = element_text(size = 8),
        legend.position = "none",
        axis.text.x=element_text(angle=-90),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank())+
  facet_wrap(~crime_type)
```
<br>
 
```{r}

amdata_tbl %>% 
  ggplot() +
  geom_col(aes(x = crime_type, y = pct_agg, fill = crime_type))+
  theme_minimal()+
  labs(title = "Portion of Aggravating Role Adjustments by Crime Type and Demographic")+
  theme(title = element_text(size = 8),
        legend.position = "none",
        axis.text.x=element_text(angle=-90),
        axis.text.y = element_text(size = 6),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.title = element_blank())+
  scale_fill_brewer(palette = "Dark2")+
  facet_wrap(~racesex_clean, ncol = 2)

```
