---
title: "Demographic Sentencing Disparities Baseline Switch"
output: word_document
date: "October 29, 2022"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 14, fig.height = 8)

library(tidyverse)
library(flextable)
library(janitor)
library(scales)

data <- read_csv(here::here("data/io.csv"))

postprotect <- data %>% 
  filter((!is.na(sentdate) & "2003-05-01"<=sentdate) | (!is.na(sentmonyr) & sentmonyr<"2004-07-01")) #06-24-04

postbooker <- data %>% 
  filter("2005-01-01"<=sentmonyr & sentmonyr<="2007-12-01" & (BOOKPOST!=0 | is.na(BOOKPOST))) #dates are not right #12-10-2007

postgall <- data %>% 
  filter("2007-12-01"<=sentmonyr & sentmonyr<="2011-09-01")

postreport <- data %>% 
  filter("2011-10-01"<=sentmonyr & sentmonyr <="2016-09-01")


### add in more recent data

```

## White Male Baseline Regressions

```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```

### Post-Protect

```{r}
model2_factor_protect <- model2(postprotect)

as_flextable(model2_factor_protect)
```
<br>
<br>

### Post-Booker

```{r}
model2_factor_booker <- model2(postbooker)

as_flextable(model2_factor_booker)
```

```{r, include = F}
summary(model2_factor_booker)
```
<br>
<br>

### Post-Gall

```{r}
model2_factor_gall <- model2(postgall)

as_flextable(model2_factor_gall)
```

```{r, include = F}
summary(model2_factor_gall)
```
<br>
<br>

### Post-Report

```{r}
model2_factor_report <- model2(postreport)

as_flextable(model2_factor_report)
```
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
<br>
  
\newpage 
 
 
## Demographic Disparities in Sentencing: Bar Graphs

```{r}
logtrans <- function(x) {
  round((exp(as.numeric(x))-1)*100, 1)
}

demogbar <- function(df){
  df %>% 
    ggplot(aes(x = period, y = value, fill = period))+
    geom_bar(stat="identity")+
    geom_text(aes(label=value), vjust=0.3, size = 6)+ 
    labs(y = "Percent", x = "Period",
         caption = "Values not statistically significant at the 0.01 level are omitted.")+
    theme_minimal() + 
    facet_wrap(~demog)+
    scale_fill_brewer(type = "qual", palette = 6) +
    theme(strip.text = element_text(size = 25),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 22),
          axis.title.y = element_text(size = 22),
          axis.title.x = element_text(size = 22),
          plot.caption = element_text(size = 20),
          axis.text.y = element_blank(),
          axis.text.x=element_blank())
  
}

```


### White Male Baseline

```{r}

extract_coeffs <- function(df){
  summary(df)[[4]] %>% 
    data.frame() %>% 
    tibble::rownames_to_column("var") %>% 
    clean_names() %>% 
    select(var, estimate, pr_t) %>% 
    filter(str_detect(var, "male")) %>% 
    mutate(
      coeff = case_when(
        pr_t<=0.01 ~ estimate,
        pr_t>0.01 ~ NA_real_,
        TRUE ~ NA_real_),
      value = logtrans(coeff),
      demog = case_when(
        str_detect(var, "whitemale") ~ "White Male",
        str_detect(var, "whitefemale") ~ "White Female",
        str_detect(var, "blackmale") ~ "Black Male",
        str_detect(var, "blackfemale") ~ "Black Female",
        str_detect(var, "hispmale") ~ "Hispanic Male",
        str_detect(var, "hispfemale") ~ "Hispanic Female",
        str_detect(var, "othermale") ~ "Other Male",
        str_detect(var, "otherfemale") ~ "Other Female"
      )) %>% 
    select(demog, value)
}

protect_wm <- extract_coeffs(model2_factor_protect) %>% mutate(period = "Protect")
booker_wm <- extract_coeffs(model2_factor_booker) %>% mutate(period = "Booker")
gall_wm <- extract_coeffs(model2_factor_gall) %>% mutate(period = "Gall")
report_wm <- extract_coeffs(model2_factor_report) %>% mutate(period = "Post-Report")


bar_data_wm <- bind_rows(protect_wm,
                         booker_wm,
                         gall_wm,
                         report_wm) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report"))
```

```{r}
bar_data_wm %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_wm %>% filter(str_detect(demog, "Male")) %>% demogbar()

#add in statistical significance dots
#fix position of bar labels

```
<br>
<br>

### White Female Baseline

```{r}
model2_wf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 7) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_protect_wf <- model2_wf(postprotect)
model2_booker_wf <- model2_wf(postbooker)
model2_gall_wf <- model2_wf(postgall)
model2_report_wf <- model2_wf(postreport)

```

```{r}
protect_wf <- extract_coeffs(model2_protect_wf) %>% mutate(period = "Protect")
booker_wf <- extract_coeffs(model2_booker_wf) %>% mutate(period = "Booker")
gall_wf <- extract_coeffs(model2_gall_wf) %>% mutate(period = "Gall")
report_wf <- extract_coeffs(model2_report_wf) %>% mutate(period = "Post-Report")


bar_data_wf <- bind_rows(protect_wf,
                         booker_wf,
                         gall_wf,
                         report_wf) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report"))

bar_data_wf %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_wf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
<br>
<br>

### Hispanic Female Baseline

```{r}
model2_hf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 3) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_protect_hf <- model2_hf(postprotect)
model2_booker_hf <- model2_hf(postbooker)
model2_gall_hf <- model2_hf(postgall)
model2_report_hf <- model2_hf(postreport)
```

```{r}
protect_hf <- extract_coeffs(model2_protect_hf) %>% mutate(period = "Protect")
booker_hf <- extract_coeffs(model2_booker_hf) %>% mutate(period = "Booker")
gall_hf <- extract_coeffs(model2_gall_hf) %>% mutate(period = "Gall")
report_hf <- extract_coeffs(model2_report_hf) %>% mutate(period = "Post-Report")


bar_data_hf <- bind_rows(protect_hf,
                         booker_hf,
                         gall_hf,
                         report_hf) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report"))

bar_data_hf %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_hf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
<br>
<br>


### Black Female Baseline

```{r}
model2_bf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 1) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_protect_bf <- model2_bf(postprotect)
model2_booker_bf <- model2_bf(postbooker)
model2_gall_bf <- model2_bf(postgall)
model2_report_bf <- model2_bf(postreport)
```

```{r}
protect_bf <- extract_coeffs(model2_protect_bf) %>% mutate(period = "Protect")
booker_bf <- extract_coeffs(model2_booker_bf) %>% mutate(period = "Booker")
gall_bf <- extract_coeffs(model2_gall_bf) %>% mutate(period = "Gall")
report_bf <- extract_coeffs(model2_report_bf) %>% mutate(period = "Post-Report")


bar_data_bf <- bind_rows(protect_bf,
                         booker_bf,
                         gall_bf,
                         report_bf) %>% 
  mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report"))

bar_data_bf %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_bf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

\newpage

## Descriptive Tables

```{r}
#select(SENSPLT0, racesex, violent, sexual, sexual2, porn, drugtraff, othdrug, whitecoll, immigration, drug, othtype, othtype2)

crimetypetable_base <- function(df){
  df %>% 
    select(SENSPLT0, racesex, violent, sexual2, porn, drugtraff, whitecoll, immigration, othtype2) %>% 
    pivot_longer(violent:othtype2, names_to = "crime_type", values_to = "truefalse") %>% 
    filter(truefalse==TRUE & !is.na(racesex)) %>% 
    select(-truefalse) %>% 
    group_by(racesex, crime_type) %>% 
    summarize(n = n(),
              avg_sent = round(mean(SENSPLT0, na.rm = T), 1)) %>% 
    mutate(racesex = case_when(
        str_detect(racesex, "whitemale") ~ "White Male",
        str_detect(racesex, "whitefemale") ~ "White Female",
        str_detect(racesex, "blackmale") ~ "Black Male",
        str_detect(racesex, "blackfemale") ~ "Black Female",
        str_detect(racesex, "hispmale") ~ "Hispanic Male",
        str_detect(racesex, "hispfemale") ~ "Hispanic Female",
        str_detect(racesex, "othermale") ~ "Other Male",
        str_detect(racesex, "otherfemale") ~ "Other Female"),
          crime_type = crime_type %>% 
          str_replace("drugtraff", "Drug Trafficking") %>% 
          str_replace("whitecoll", "White Collar") %>% 
          str_replace("othtype2", "Other Types") %>% 
          str_replace("sexual2", "Sexual") %>% 
          str_to_title()) %>% 
    rename("Crime Type" = "crime_type")
}


crimetypetable_n <- function(df){
  df %>% 
    crimetypetable_base() %>% 
    select(-avg_sent) %>% 
    pivot_wider(names_from = racesex, values_from = n) %>% 
    adorn_totals(where = c("row", "col")) %>% 
    flextable() %>% 
    fontsize(i = NULL, j = NULL, size = 10, part = "body") %>% 
    fontsize(i = NULL, j = NULL, size = 10, part = "header") %>% 
    set_caption("Total Number of Offenders for Each Demographic Group and Crime Type") %>% 
    set_table_properties(width = .7, layout = "autofit")
}

crimetypetable_pct <- function(df){
  df %>% 
    crimetypetable_base() %>% 
    select(-avg_sent) %>% 
    pivot_wider(names_from = racesex, values_from = n) %>% 
    adorn_percentages("col") %>% 
    mutate(across('Black Female':'White Male', ~ scales::percent(round(., 3)))) %>% 
    flextable() %>% 
    fontsize(i = NULL, j = NULL, size = 10, part = "body") %>% 
    fontsize(i = NULL, j = NULL, size = 10, part = "header") %>% 
    set_caption("Offenders by Demographic Group and Crime Type: Percentage of Demographic Group") %>% 
    set_table_properties(width = .7, layout = "autofit")
}

crimetypetable_sent <- function(df){
  df %>% 
    crimetypetable_base() %>% 
    select(-n) %>% 
    pivot_wider(names_from = racesex, values_from = avg_sent) %>% 
    flextable() %>% 
    fontsize(i = NULL, j = NULL, size = 10, part = "body") %>% 
    fontsize(i = NULL, j = NULL, size = 10, part = "header") %>% 
    set_caption("Average Sentence Length in Months, by Demographic Group and Crime Type") %>% 
    set_table_properties(width = .7, layout = "autofit")
}

model2filter <- function(df){
  df %>% 
    select(SENSPLT0,
           logsplit,
           logmin,
           violent,
           drugtraff,
           sexual2,
           porn,
           immigration,
           othtype2,
           whitecoll,
           upward,
           downdep,
           subasst,
           mandmin2,
           NEWCNVTN,
           custody,
           racesex,
           agedummy,
           educ,
           citizen) %>% 
    filter(if_all(everything(), ~ !is.na(.)))
    
}
```

**Total Number of Individual Offenders and Average Sentence Length, by Demographic Group and Type of Crime**

### Post-Protect

```{r}
postprotect %>% model2filter() %>%  crimetypetable_n()
postprotect %>% model2filter() %>%  crimetypetable_pct()
postprotect %>% model2filter() %>% crimetypetable_sent()
```

\newpage

### Post-Booker

```{r}
postbooker %>% model2filter() %>% crimetypetable_n()
postbooker %>% model2filter() %>%  crimetypetable_pct()
postbooker %>% model2filter() %>% crimetypetable_sent()
```

\newpage

### Post-Gall

```{r}
postgall %>% model2filter() %>% crimetypetable_n()
postgall %>% model2filter() %>%  crimetypetable_pct()
postgall %>% model2filter() %>% crimetypetable_sent()
```

\newpage

### Post-Report

```{r}
postreport %>% model2filter() %>% crimetypetable_n()
postreport %>% model2filter() %>%  crimetypetable_pct()
postreport %>% model2filter() %>% crimetypetable_sent()
```



