---
title: "Region Breakdown Pt 3 - Female"
output: html_document
date: "2023-10-31"
---

# Introduction 
This section of the region breakdown isolates only female identifying defendants. Otherwise, this structure is the exact same as Part 1 and provides only descriptive statistics. 


# Location 


```{r, include=FALSE}

library(tidyverse)
library(flextable)
library(janitor)
library(scales)
library(readxl)

data <- read_csv(here::here("data/io.csv"))
district <- data %>% 
  filter(districtName %in% c("Texas North", "Texas West", "Texas South", "Texas East","Louisiana East", "Louisiana West", "New Mexico", "Florida North", "Florida Middle", "Florida South")) %>% 
  filter(str_detect(racesex_clean, 'Female'))

postprotect <- district %>% 
  filter(postprotect) #06-24-04

postbooker <- district %>% 
  filter(postbooker) #dates are not right #12-10-2007

postgall <- district %>% 
  filter(postgall)

postreport <- district %>% 
  filter(postreport)

present <- district %>% 
  filter(present)


highlight_above_avg <- function(ft){
  ft %>% 
     #bg(i = ~ `Black Male`>Overall, 
      #j = ~ `Black Male`, 
      #bg="#Fba894") %>% 
     #bg(i = ~ `Hispanic Male`>Overall, 
      #j = ~ `Hispanic Male`, 
     #bg="#Fba894") %>% 
     #bg(i = ~ `White Male`>Overall, 
        #j = ~ `White Male`, 
        #bg="#Fba894") %>% 
     #bg(i = ~ `Other Male`>Overall, 
        #j = ~ `Other Male`, 
        #bg="#Fba894") %>% 
    bg(i = ~ `Black Female`>Overall, 
       j = ~ `Black Female`, 
       bg="#Fba894") %>% 
    bg(i = ~ `Hispanic Female`>Overall, 
       j = ~ `Hispanic Female`, 
       bg="#Fba894") %>% 
    bg(i = ~ `White Female`>Overall, 
       j = ~ `White Female`, 
       bg="#Fba894") %>% 
    bg(i = ~ `Other Female`>Overall, 
       j = ~ `Other Female`, 
       bg="#Fba894")
}
```

\newpage

## Distribution of Sentences 
```{r}
# ggplot(data = district, aes(x = districtName, y = SENSPLT0)) +
#   geom_boxplot() 

boxplot(SENSPLT0 ~ districtName, data = district, 
        outline = FALSE, las = 2, ylab = "", xlab = '', cex.axis = .7)#does not include outliers 

```
*H0: All boxes would be of equal widths and have the same average sentence lengths. This means that the black bars would be all aligned, and the ranges would also be exactly the same.*
*H1: The boxes would not be of equal widths with evenly aligned black bars.* 

From the above visualization, we can note the following observations: 
1. Northern and Eastern Texas have larger average sentences for women

\newpage

## Average Sentence, by Federal District and Period 

```{r full picture}
dem_avg <- function(df){
  df %>%
    group_by(districtName) %>%
    summarize(avg_sent = mean(SENSPLT0, na.rm=T))
}
list_periods <- list(postprotect, postbooker, postgall, postreport, present)

avg_sent_dem_tbl <- map(list_periods, dem_avg) %>%
  bind_rows(.id = "Period") %>%
  mutate(avg_sent = round(avg_sent),
         Period = recode(as.character(Period),
                         "1" = "Protect", "2" = "Booker", "3" = "Gall", "4" = "Post-Report", "5" = "Present")) %>%
  pivot_wider(names_from = Period, values_from = avg_sent) %>%
  rename("Group" = districtName)

avg_sent_dem_tbl

#--- visualization 

df <- avg_sent_dem_tbl %>% 
  pivot_longer(!Group, names_to = 'period', values_to= 'avg_sent')

df$num_period <- as.numeric(factor(df$period, levels = c("Protect", "Booker", "Gall", "Post-Report", "Present")))


ggplot(data =df, aes(x = num_period, y = avg_sent, col = Group)) + 
  geom_point() + 
  geom_line() + 
  scale_x_continuous(labels = c("Protect", "Booker", "Gall", "Post-Report", "Present")) + 
  xlab('Period') + 
  ylab('Avg Sentence Length')
```
*H0: There would be no distinguishing lines and/or trends between each district as average sentence lengths across districts should have been the same in each period.*
*H1: There are distinguishing lines and/or trends between each district.*

1. Over time, average sentence length for women in Northern and Eastern Texas went up after Booker
2. Average sentence length for women in Western Louisiana went up after the Post-Report period. 

Because of this trend increase, I was interested to see the breakdown between demographics of women. 


```{r}
dem_avg2 <- function(df){
  df %>%
    group_by(districtName, racesex_clean) %>%
    summarize(avg_sent = mean(SENSPLT0, na.rm=T))
}


list_periods <- list(postprotect, postbooker, postgall, postreport, present)

avg_sent_dem_tbl2 <- map(list_periods, dem_avg2) %>%
  bind_rows(.id = "Period") %>%
  mutate(avg_sent = round(avg_sent),
         Period = recode(as.character(Period),
                         "1" = "Protect", "2" = "Booker", "3" = "Gall", "4" = "Post-Report", "5" = "Present")) %>%
  pivot_wider(names_from = Period, values_from = avg_sent) %>%
  rename("Group" = districtName)

avg_sent_dem_tbl2

#--- visualization 

df <- avg_sent_dem_tbl2 %>% 
  pivot_longer(!Group & !racesex_clean, names_to = 'period', values_to= 'avg_sent')

df$num_period <- as.numeric(factor(df$period, levels = c("Protect", "Booker", "Gall", "Post-Report", "Present")))

ggplot(data =df%>% filter(Group %in% c("Texas North")), aes(x = num_period, y = avg_sent, col = racesex_clean)) + 
  geom_point() + 
  geom_line() + 
  scale_x_continuous(labels = c("Protect", "Booker", "Gall", "Post-Report", "Present")) + 
  xlab('Period') + 
  ylab('Avg Sentence Length') + 
  ggtitle("Avg Sentence Length Over Time in North Texas")


ggplot(data =df%>% filter(Group %in% c("Texas East")), aes(x = num_period, y = avg_sent, col = racesex_clean)) + 
  geom_point() + 
  geom_line() + 
  scale_x_continuous(labels = c("Protect", "Booker", "Gall", "Post-Report", "Present")) + 
  xlab('Period') + 
  ylab('Avg Sentence Length') + 
  ggtitle("Avg Sentence Length Over Time in East Texas")


ggplot(data =df%>% filter(Group %in% c("Louisiana West")), aes(x = num_period, y = avg_sent, col = racesex_clean)) + 
  geom_point() + 
  geom_line() + 
  scale_x_continuous(labels = c("Protect", "Booker", "Gall", "Post-Report", "Present")) + 
  xlab('Period') + 
  ylab('Avg Sentence Length') + 
  ggtitle("Avg Sentence Length Over Time in Lousiana West")
```

1. White Females make some of the largest upward trends among other female demographic groups. 
2. There are general upward trends in average sentence length for women in these three districts. 


\newpage 

## Average Sentence Length 
```{r demographic }
df <- district%>%
  select(SENSPLT0, districtName, racesex_clean) %>%
  group_by(districtName, racesex_clean) %>%
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>%
  ungroup() %>%
  pivot_wider(names_from = racesex_clean, values_from = avg_sent)
df %>% flextable()

```

In this table, we present the numerical values for average sentence length spread out across district and demographic. Below, we visualize this information to better understand from a top down perspective what this table describes. 


## Average Sentence Length in each District

```{r}

par(mfrow=c(2,2))
# Loop through rows and create bar charts
for (i in 1:nrow(df)) {
  # Select data for the current row
  current_row <- df[i, ]

  # Extract the fedNum and remove it from the data for plotting
  name <- current_row$districtName
  current_row <- current_row[-1]

  # Create a bar chart for the current row
  barplot(as.matrix(current_row), names.arg = names(current_row),
          main = paste("Avg Sentence Length in", name), legend = TRUE,cex.names= 0.6,las =2 )
}
```

*H0: Average sentence length across district would be exactly the same*
*H1: Average sentence length across district are not the same*

These bar graphs show average sentence lengths all time across female demographic groups, and they appear to reflect similar observations from the line graphs. 


## Average Sentence for each Demographic 

```{r}
par(mfrow=c(2,2))
# Get the column names excluding the first column 
columns <- names(df)[-1]


# Loop through columns and create bar charts
for (col in columns) {
  # Create a bar chart for the current column
  barplot(df[[col]], names.arg = df$districtName,
          main = paste("Avg Sentence Length for", col),
          col = "steelblue", cex.names = .6, las=2)
}
```

*H0: Average sentence length across demographic are exactly the same*
*H1: Average sentence length across demographic are not the same*

1. The largest average sentence lengths for Hispanic Female, Other Female, and White Female are in Northern Texas. 

# Crime Type 

## Average Sentence by Crime Type and Federal District 
```{r}
crime_df <- district%>% 
  select(SENSPLT0,  'District Name'=districtName  , crime_type) %>% 
  group_by(`District Name`, crime_type) %>% 
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = crime_type, values_from = avg_sent) 

crime_df %>% flextable()
```

The table above is the average sentence length by crime type and district. 



```{r}

par(mfrow=c(2,2))
# Loop through rows and create bar charts
for (i in 1:nrow(crime_df)) {
  # Select data for the current row
  current_row <- crime_df[i, ]

  # Extract the fedNum and remove it from the data for plotting
  name <- current_row$`District Name`
  current_row <- current_row[-1]

  # Create a bar chart for the current row
  barplot(as.matrix(current_row), names.arg = names(current_row),
          ylab = "Avg Sentence Length in Months", 
          main = paste("Avg Sentence Length in", name), legend = TRUE,cex.names= 0.6,las =2 )
}
```

*H0: The average sentence length for each crime type remains consistent across all districts*
*H1: The average sentence length for each crime type is not consistent across all districts*

1. Despite that Porn has the largest average sentence length in seven of the nine districts, Eastern Louisiana has an incredibly small sentence lengths for porn sentences for women. Western Lousiana also has an incredibly small sentence length for porn. 

2. The largest average sentence length for women in Eastern Louisiana is Drug Trafficking, whereas in Western Louisiana, it is sexual. 


## Average Sentence for each Demographic 

```{r}
par(mfrow=c(2,2))
# Get the column names excluding the first column 
columns <- names(crime_df)[-1]
# Loop through columns and create bar charts
for (col in columns) {
  # Create a bar chart for the current column
  barplot(crime_df[[col]], names.arg = crime_df$`District Name`,
          main = paste("Avg Sentence Length for", col), ylab = "Avg Sentence in Months", 
          col = "goldenrod", cex.names = .6, las=2)
}
```

*H0: The average sentence length for each crime type remains consistent across all districts*
*H1: The average sentence length for each crime type is not consistent across all districts*

1. For immigration based offenses, Northern Texas had higher average sentences than most other districts despite the fact that immigration in Northern Texas did not have the largest average sentence length. 


## Drug Trafficking Sentence Length

```{r}
par(mfrow = c(2,2))
drug <-  present %>% 
  filter(crime_type== "Drug Trafficking") %>% 
  select(crime_type, racesex_clean, districtName, SENSPLT0) %>% 
  group_by(racesex_clean, districtName) %>% 
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = districtName, values_from = avg_sent)


# Get the column names excluding the first column (fedNum)
columns <- names(drug)[-1]
# Loop through columns and create bar charts
for (col in columns) {
  # Create a bar chart for the current column
  barplot(drug[[col]], names.arg = drug$racesex_clean,
          main = paste("Drug Trafficking in", col), ylab = 'Avg Sentence in Months',
          col = "cornflowerblue", cex.names = .7, las = 2)
}
```

*H0: Sentencing lengths for drug trafficking offenses are the same across all demographics and districts.*
*H1: Sentencing lengths for drug trafficking offenses are not the same across all demographics and districts.* 

There aren't many outstanding notes for this passage. 

## Immigration Sentence Length

```{r}
par(mfrow = c(2,2))
imm <-  present %>% 
  filter(crime_type== "Immigration") %>% 
  select(crime_type, racesex_clean, districtName, SENSPLT0) %>% 
  group_by(racesex_clean, districtName) %>% 
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = districtName, values_from = avg_sent)


# Get the column names excluding the first column (fedNum)
columns <- names(imm)[-1]
# Loop through columns and create bar charts
for (col in columns) {
  # Create a bar chart for the current column
  barplot(imm[[col]], names.arg = imm$racesex_clean,
          main = paste("Bar Chart for", col), ylab = 'Avg Sentence in Months',
          col = "coral", cex.names = .6, las = 2)
}
```

*H0: Sentencing lengths for immigration offenses are the same across all demographics and districts.*
*H1: Sentencing lengths for immigration offenses are not the same across all demographics and districts.* 

1. White and Hispanic females comprise most immigration based cases. 


\newpage

## Detailed Sentence Length Breakdown, by Guideline and Demographic Group

The table below presents a detailed sentence length breakdown for each guideline 
```{r}
#crime names
crime_name_tbl <- read_excel(here::here("data/Type of Crime.xlsx")) %>% 
  clean_names() %>% 
  select(statute, offense) %>% 
  mutate(offense = offense %>% 
           str_to_title() %>% 
           str_trunc(20, "right"))

district %>% 
  select(districtName, crime_type, racesex_clean, GDLINEHI, GLMIN) %>% 
  group_by(districtName, crime_type, GDLINEHI) %>% 
  mutate(gdline_avg = round(mean(GLMIN, na.rm = T))) %>% 
  group_by(crime_type, gdline_avg, racesex_clean, GDLINEHI) %>% 
  summarize(glmin_avg = round(mean(GLMIN, na.rm = T))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = racesex_clean, values_from = glmin_avg) %>% 
  select("Crime Type" = crime_type, Guideline = GDLINEHI, contains(" Male"), contains(" Female"), Overall = gdline_avg) %>% 
  arrange(Guideline) %>% 
  left_join(crime_name_tbl, by = c("Guideline" = "statute")) %>% 
  select(`Crime Type`, Guideline, Offense = offense, everything()) %>% 
  flextable() %>% 
  set_caption("Average Sentence Length in Months, by Demographic Group and Guideline, FY 2002 - 2021") %>% 
  fontsize(i = NULL, j = NULL, size = 8.5, part = "body") %>%
  fontsize(i = NULL, j = NULL, size = 8.5, part = "header") %>%
  highlight_above_avg() %>% 
  set_table_properties(width = 1, layout = "autofit")
```

