---
title: "Region Breakdown"
output: html_document
date: "2023-10-31"
---


```{r setup, include=FALSE}

library(tidyverse)
library(flextable)
library(janitor)
library(scales)
library(readxl)

data <- read_csv(here::here("data/io.csv"))
district <- data %>% filter(districtName %in% c("Texas North", "Texas West", "Texas South", "Texas East","Louisiana East", "Louisiana West", "New Mexico", "Florida North", "Florida Middle", "Florida South"))

femDisCrime <- district %>% 
  select(districtName, crime_type, racesex_clean, GLMIN, SENSPLT0) %>%
  filter(str_detect(racesex_clean, 'Female')) %>% 
  mutate(crime_type = if_else(is.na(crime_type),'NA', crime_type))

postprotect <- district %>% 
  filter(postprotect) #06-24-04

postbooker <- district %>% 
  filter(postbooker) #dates are not right #12-10-2007

postgall <- district %>% 
  filter(postgall)

postreport <- district %>% 
  filter(postreport)

present <- district %>% 
  filter(present)

highlight_above_avg <- function(ft){
  ft %>% 
    # bg(i = ~ `Black Male`>Overall, 
    #  j = ~ `Black Male`, 
    #  bg="#Fba894") %>% 
    # bg(i = ~ `Hispanic Male`>Overall, 
    #    j = ~ `Hispanic Male`, 
    #    bg="#Fba894") %>% 
    # bg(i = ~ `White Male`>Overall, 
    #    j = ~ `White Male`, 
    #    bg="#Fba894") %>% 
    # bg(i = ~ `Other Male`>Overall, 
    #    j = ~ `Other Male`, 
    #    bg="#Fba894") %>% 
    bg(i = ~ `Black Female`>Overall, 
       j = ~ `Black Female`, 
       bg="#Fba894") %>% 
    bg(i = ~ `Hispanic Female`>Overall, 
       j = ~ `Hispanic Female`, 
       bg="#Fba894") %>% 
    bg(i = ~ `White Female`>Overall, 
       j = ~ `White Female`, 
       bg="#Fba894") %>% 
    bg(i = ~ `Other Female`>Overall, 
       j = ~ `Other Female`, 
       bg="#Fba894")
}
```


## Distribution of Sentences 
```{r}
#only for females

ggplot(data = femDisCrime, aes(x = districtName, y = SENSPLT0)) + 
  geom_boxplot()

```



## Average Sentence, by Federal District and Period 

```{r full picture}
dem_avg <- function(df){
  df %>%
    group_by(districtName, racesex_clean) %>%
    summarize(avg_sent = mean(SENSPLT0, na.rm=T))
}

list_periods <- list(postprotect, postbooker, postgall, postreport, present)

avg_sent_dem_tbl <- map(list_periods, dem_avg) %>%
  bind_rows(.id = "Period") %>%
  mutate(avg_sent = round(avg_sent),
         Period = recode(as.character(Period),
                         "1" = "Protect", "2" = "Booker", "3" = "Gall", "4" = "Post-Report", "5" = "Present")) %>%
  pivot_wider(names_from = Period, values_from = avg_sent) %>%
  rename("Group" = districtName)

avg_sent_dem_tbl


```
- Florida North average sentences went down a lot over time 
- Texas East and Texas North went up 
- Texas West went down 

## Average Sentence by Demographic and Federal District

```{r demographic }


par(mfrow=c(2,2))
df <- femDisCrime%>%
  select(SENSPLT0, districtName, racesex_clean) %>%
  group_by(districtName, racesex_clean) %>%
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>%
  ungroup() %>%
  pivot_wider(names_from = racesex_clean, values_from = avg_sent)

df %>% flextable()


# Loop through rows and create bar charts
for (i in 1:nrow(df)) {
  # Select data for the current row
  current_row <- df[i, ]

  # Extract the fedNum and remove it from the data for plotting
  name <- current_row$districtName
  current_row <- current_row[-1]

  # Create a bar chart for the current row
  barplot(as.matrix(current_row), names.arg = names(current_row),
          main = paste("Bar Chart for ", name), legend = TRUE,cex.names= 0.6,las =2 )
}

# Get the column names excluding the first column 
columns <- names(df)[-1]


# Loop through columns and create bar charts
for (col in columns) {
  # Create a bar chart for the current column
  barplot(df[[col]], names.arg = df$districtName,
          main = paste("Bar Chart for", col),
          col = "steelblue", cex.names = .8, las = 2)
}

```

- Texas East and Texas North have much higher avg sentences for Hispanic populations. 
- New Mexico has a much higher avg sentence for Other Male 

# Crime Type 

## Average Sentence by Crime Type and Federal District 
```{r}
crime_df <- 
  femDisCrime%>% 
  group_by(districtName, crime_type, racesex_clean) %>%
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>% 
  group_by(districtName, crime_type, avg_sent, racesex_clean) %>% 
  ungroup() %>% 
  #pivot_wider(names_from = racesex_clean, values_from = avg_sent) %>% 
  select("Race/Sex" = racesex_clean,  Overall = avg_sent) %>% 
  flextable() %>% 
  highlight_above_avg()

crime_df %>% flextable()

```

- Largest drug trafficking offenses in Florida North, followed by Texas North. 

```{r}
par(mfrow = c(2,2))
drug <-  present %>% 
  filter(crime_type== "Drug Trafficking") %>% 
  select(crime_type, racesex_clean, districtName, SENSPLT0) %>% 
  group_by(racesex_clean, districtName) %>% 
  summarize(avg_sent = round(mean(SENSPLT0, na.rm = T))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = districtName, values_from = avg_sent)



# Get the column names excluding the first column (fedNum)
columns <- names(drug)[-1]
# Loop through columns and create bar charts
for (col in columns) {
  # Create a bar chart for the current column
  barplot(drug[[col]], names.arg = drug$racesex_clean,
          main = paste("Bar Chart for", col), ylab = col,
          col = "cornflowerblue", cex.names = .7, las = 2)
}
```
- Hispanic Female sentence length almost the same as Hispanic Male in Louisiana West 
- Hispanic Female higher than Black Male in Texas North and almost as high as Hispanic Male



## Detailed Sentence Length Breakdown, by Guideline and Demographic Group

```{r}

# disCrime <- district %>% 
#   # select(districtName, crime_type, racesex_clean, GLMIN, SENSPLT0) %>% 
#   # group_by(districtName, crime_type) %>% 
#   # summarize(gdline_avg = round(mean(GLMIN, na.rm = T)), 
#   #        glmin_avg = round(mean(GLMIN, na.rm = T)), 
#   #        sent_avg = round(mean(SENSPLT0, na.rm = T))) %>% ungroup()
#   select(districtName, crime_type, racesex_clean, SENSPLT0) %>% 
#   group_by(districtName, crime_type, racesex_clean) %>% 
#   mutate(sent_avg = round(mean(SENSPLT0, na.rm = T)))
# disCrime
```

Two-way ANOVA: used to evaluate the effects of 2 categorical variables (and their potential interaction) on a quantitative continuous variable


```{r}
summary(disCrime)

mod <- aov(sent_avg~districtName*crime_type, data = disCrime)
library(car)

qqPlot(mod$residuals,
  id = FALSE # remove point identification
)


```


#ANCOVA
- assumes normality of residuals 
- concerned with studying regression in a set of groups
- applied in experimental studies when researchers want to remove the effects of some antecedent variable 
- combines principles of ANOVA and principles of regression 
- reduce the within-group (error) variances, making the test more efficient (powerful). ie. statistically remove the variance 

#ANOVA 
- assumes data are independent 
- assumes variance is the same 
- assumes normal distribution 

# Three Way ANOVA 
assesses whether there is an interaction effect between three independent categorical variables on a continuous outcome variabel 
```{r}
# model <- lm(sent_avg ~ districtName + racesex_clean + crime_type + 
#               districtName:racesex_clean + racesex_clean:crime_type + districtName:crime_type, data = disCrime)
# model



library(rstatix)
district %>% 
  group_by(districtName, crime_type, racesex_clean) %>% 
  get_summary_stats(SENSPLT0, type = 'mean_sd')



ggplot(data = femDisCrime, mapping =aes(x=racesex_clean, y = 'SENSPLT0')) + 
  geom_boxplot()

library(ggpubr)
#visualization 
ggboxplot(
  femDisCrime, x = 'districtName', y = 'SENSPLT0',
  color = 'racesex_clean', 
  facet.by  = 'crime_type'
)
```



```{r}
#crime names
crime_name_tbl <- read_excel(here::here("data/Type of Crime.xlsx")) %>% 
  clean_names() %>% 
  select(statute, offense) %>% 
  mutate(offense = offense %>% 
           str_to_title() %>% 
           str_trunc(20, "right"))

district %>% 
  select(districtName, crime_type, racesex_clean, GDLINEHI, GLMIN) %>% 
  group_by(districtName, crime_type, GDLINEHI) %>% 
  mutate(gdline_avg = round(mean(GLMIN, na.rm = T))) %>% 
  group_by(crime_type, gdline_avg, racesex_clean, GDLINEHI) %>% 
  summarize(glmin_avg = round(mean(GLMIN, na.rm = T))) %>% 
  ungroup() %>% 
  pivot_wider(names_from = racesex_clean, values_from = glmin_avg) %>% 
  select("Crime Type" = crime_type, Guideline = GDLINEHI, contains(" Male"), contains(" Female"), Overall = gdline_avg) %>% 
  arrange(Guideline) %>% 
  left_join(crime_name_tbl, by = c("Guideline" = "statute")) %>% 
  select(`Crime Type`, Guideline, Offense = offense, everything()) %>% 
  flextable() %>% 
  set_caption("Average Sentence Length in Months, by Demographic Group and Guideline, FY 2002 - 2021") %>% 
  fontsize(i = NULL, j = NULL, size = 8.5, part = "body") %>%
  fontsize(i = NULL, j = NULL, size = 8.5, part = "header") %>%
  highlight_above_avg() %>% 
  set_table_properties(width = 1, layout = "autofit")
```


# Gender Baseline

Rodriguez, Curry, and Lee(2006) found that women overall were less likely to be sentenced and if so, received shorted sentences than men. Goulette, Wooldredge, and Lawrence(2015) found that the intersection of race and gender impacted Bond amounts, and minority females received higher bond amounts than White women, and specific crimes (i.e."rape, robbery, burglary, theft, fraud, drugs, and domestic violence/nonsupport") significantly impacted bond amounts as well. 

Because of the length of this, I have a second document which is a copy of the demographic sentencing disparities based on only districts. This is region_demographic2.Rmd.    