---
title: "Demographic Baseline 2017 2021"
fontsize: 12pt
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    toc_depth: 5
    number_sections: true
  pdf_document: default
urlcolor: magenta
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```   


```{r}
library(bigrquery)
library(DBI)
library(tidyverse)
library(flextable)
library(janitor)
library(scales)
```
-- SENTRANGE replaces DEPART 2018 onwards 
-- IMMIGRANT and OTHTYPE look in codebook 
-- supplemental datafiles on the commission's website for the more recent data 

# Step 1: Making another io_raw.csv 


```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "balmy-coral-330818",
  dataset = "bqtest"
  )

dbListTables(con)

```


## 2017 Raw Data 
In fy17_1: 
SENSPLT0, GLMIN, TOTCHPTS, IS924C, WEAPSOC, STATMIN, CAROFFAP, ACCAP, SAFE, NEWCNVTN, PRESENT, MITROLHI, AGGROLHI, NEWRACE, MONSEX, AGE, EDUCATN, NEWCIT

In fy17_2: 
GDLINEHI

```{r}
query2017_1 = dbSendQuery(con, "select `id`,`SENTMON`,`SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `BOOKERCD`from fy17_1")
data2017_1 <-  dbFetch(query2017_1)
head(data2017_1)

query2017_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy17_2 ")
data2017_2 = dbFetch(query2017_2)

head(data2017_2)

data2017 <- merge(data2017_1, data2017_2, by = 'id')
data2017$opafy <- 2017

data2017 <- subset(data2017, select= -c(1))
head(data2017)

```

## 2018
In opafy18nid_1: 
SENSPLT0, GLMIN, TOTCHPTS, IS924C, WEAPSOC, STATMIN, CAROFFAP, ACCAP, SAFE, NEWCNVTN, PRESENT, MITROLHI, AGGROLHI, NEWRACE, MONSEX, AGE, EDUCATN, NEWCIT

In opafy18nid_3: 
GDLINEHI
```{r}
query2018_1 = dbSendQuery(con, "select `id`,`sentmon`, `sentyr`,`sensplt0`, `glmin`,`totchpts`,`is924c`,`weapsoc`, `statmin`, `caroffap`,`accap`, `safe`, `newcnvtn`, `present`, `mitrolhi`, `aggrolhi`, `newrace`, `monsex`, `age`, `educatn`, `newcit` from fy18_1")

data2018_1 <-  dbFetch(query2018_1)
head(data2018_1)

query2018_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy18_3")
data2018_2 <- dbFetch(query2018_2)
head(data2018_2)

data2018 <- merge(data2018_1, data2018_2, by = 'id')

names(data2018) <- toupper(names(data2018))

data2018$opafy <- 2018
data2018 <- subset(data2018, select= -c(1))
head(data2018)
```

## 2019 
```{r}
query2019_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT` from fy19_1")

data2019_1 <-  dbFetch(query2019_1)
head(data2019_1)

query2019_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy19_2 ")
data2019_2 <- dbFetch(query2019_2)
head(data2019_2)

data2019 <- merge(data2019_1, data2019_2, by = 'id')
data2019$opafy <- 2019
data2019 <- subset(data2019, select= -c(1))
head(data2019)
```

## 2020 
```{r}
query2020_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT` from fy20_1")

data2020_1 <-  dbFetch(query2020_1)
head(data2020_1)

query2020_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy20_2 ")
data2020_2 <- dbFetch(query2020_2)
head(data2020_2)

data2020 <- merge(data2020_1, data2020_2, by = 'id')
data2020$opafy <- 2020
data2020 <- subset(data2020, select= -c(1))
head(data2020)
```


## 2021 
```{r}
query2021_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT` from fy21_1")

data2021_1 <-  dbFetch(query2021_1)
head(data2021_1)

query2021_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy21_3")
data2021_2 <- dbFetch(query2021_2)
head(data2021_2)

data2021 <- merge(data2021_1, data2021_2, by = 'id')
data2021$opafy <- 2021
data2021 <- subset(data2021, select= -c(1))
head(data2021)
```

## Merge together 
```{r}
io_raw_2017_2021 <- bind_rows(data2017, data2018, data2019, data2020, data2021)
#write.csv(io_raw_2017_2021,"~/Downloads/research/io_raw_2017_2021.csv")

io_raw_2017_2021$SENTMON <- as.numeric(io_raw_2017_2021$SENTMON)
io_raw_2017_2021$SENTYR <- as.numeric(io_raw_2017_2021$SENTYR)
io_raw_2017_2021$SENSPLT0 <- as.numeric(io_raw_2017_2021$SENSPLT0)
io_raw_2017_2021$GLMIN <-as.numeric( io_raw_2017_2021$GLMIN)
io_raw_2017_2021$TOTCHPTS <- as.numeric(io_raw_2017_2021$TOTCHPTS)
io_raw_2017_2021$IS924C <- as.numeric(io_raw_2017_2021$IS924C)
io_raw_2017_2021$WEAPSOC <- as.numeric(io_raw_2017_2021$WEAPSOC)
io_raw_2017_2021$STATMIN <- as.numeric(io_raw_2017_2021$STATMIN)
io_raw_2017_2021$CAROFFAP <- as.numeric(io_raw_2017_2021$CAROFFAP)
io_raw_2017_2021$ACCAP <- as.numeric(io_raw_2017_2021$ACCAP)
io_raw_2017_2021$SAFE <- as.numeric(io_raw_2017_2021$SAFE)#
io_raw_2017_2021$NEWCNVTN <- as.numeric(io_raw_2017_2021$NEWCNVTN)
io_raw_2017_2021$PRESENT <- as.numeric(io_raw_2017_2021$PRESENT)
io_raw_2017_2021$MITROLHI <- as.numeric(io_raw_2017_2021$MITROLHI)
io_raw_2017_2021$AGGROLHI <- as.numeric(io_raw_2017_2021$AGGROLHI)
io_raw_2017_2021$NEWRACE <- as.numeric(io_raw_2017_2021$NEWRACE)
io_raw_2017_2021$MONSEX <- as.numeric(io_raw_2017_2021$MONSEX)
io_raw_2017_2021$AGE <- as.numeric(io_raw_2017_2021$AGE)
io_raw_2017_2021$EDUCATN <- as.numeric(io_raw_2017_2021$EDUCATN)
io_raw_2017_2021$NEWCIT <- as.numeric(io_raw_2017_2021$NEWCIT)
io_raw_2017_2021$BOOKERCD <- as.numeric(io_raw_2017_2021$BOOKERCD)


str(io_raw_2017_2021)
```

# Step 2: Creating io 2017-2021 

```{r}
matchwna <- function(string, list){
  case_when(
    string %in% list ~ TRUE,
    !is.na(string) ~ FALSE,
    is.na(string) ~ NA
  )
}

#sd <- io_raw_2017_2021 %>% mutate(sentdate = paste(as.character(SENTYR),'-',as.character(SENTMON), '-01', sept = ''))

#sd$sentdate <- gsub(' ', '', sd$sentdate)
#head(sd)

data <- io_raw_2017_2021 %>% 
  mutate(SENTMON = SENTMON %>% 
           str_replace("01", "Jan") %>% 
           str_replace("02","Feb" ) %>% 
           str_replace("03","Mar") %>% 
           str_replace("04","Apr") %>% 
           str_replace("05","May") %>% 
           str_replace("06","Jun") %>% 
           str_replace("07","Jul") %>% 
           str_replace("08","Aug") %>% 
           str_replace("09","Sep") %>% 
           str_replace("10","Oct") %>% 
           str_replace("11","Nov") %>% 
           str_replace("12","Dec"), 
         logsplit = case_when(
           SENSPLT0==0.00 ~ log(0.01), 
           SENSPLT0>470 ~ log(470),
           TRUE ~ log(SENSPLT0)), 
         logmin = case_when(
           GLMIN==0.00 ~ log(0.01), 
           GLMIN>470 ~ log(470),
           TRUE ~ log(GLMIN)), 
         
         violent = matchwna(GDLINEHI, c("2A1.1", "2A1.2", "2A1.3", "2A1.4", "2A1.5", "2A2.1", "2A2.2", "2A2.3", "2A2.4", 
                                        "2A4.1", "2A4.2", "2A5.1", "2A5.2", "2A5.3", "2A6.1", "2A6.2", "2E1.3", "2E1.4", 
                                        "2E2.1", "2B3.1", "2B3.2", "2B3.3")) | str_detect(GDLINEHI, "^2K\\d{1}\\.\\d{1}$"), 
         
         sexual = matchwna(GDLINEHI, c("2A3.1", "2A3.2", "2A3.3", "2A3.4", "2G1.1", "2G1.2", "2G1.3", "2G2.1", "2G2.2", "2G2.3", "2G2.4", "2G2.5")), 
         sexual2 = matchwna(GDLINEHI, c("2A3.1", "2A3.2", "2A3.3", "2A3.4", "2G1.1", "2G1.2", "2G1.3")),
         porn = matchwna(GDLINEHI, c("2G2.1", "2G2.2", "2G2.3", "2G2.4", "2G2.5")),
         drugtraff = matchwna(GDLINEHI, c("2D1.1", "2D1.2", "2D1.5", "2D1.6", "2D1.7", "2D1.8", "2D1.9", "2D1.10", "2D1.11", "2D1.12", "2D1.13")),
         othdrug = matchwna(GDLINEHI, c("2D2.1", "2D2.2")),
         whitecoll = matchwna(GDLINEHI, c("2B1.1", "2B1.6", "2B4.1", "2B5.1", "2B5.3", "2F1.1", "2F1.2", "2R1.1", "2S1.1", "2S1.2", "2S1.3", "2S1.4")) | 
           str_detect(GDLINEHI, "^2T\\d{1}\\.\\d{1}$"), 
         
  #--------------------------------------------------------------------
         #immigration = str_detect(GDLINEHI, "^2L\\d{1}\\.\\d{1}$")) %>% 
  #mutate(postprotect = (!is.na(sentdate) & "2003-05-01"<=sentdate) | (!is.na(sentmonyr) & sentmonyr<"2004-07-01"),
         #postbooker = "2005-01-01"<=sentmonyr & sentmonyr<="2007-12-01" & (BOOKPOST!=0 | is.na(BOOKPOST)), # excluding december makes the match worse
         #postgall = "2007-12-01"<=sentmonyr & sentmonyr<="2011-09-01",
         #postreport = "2011-10-01"<=sentmonyr & sentmonyr <="2016-09-01"
 #--------------------------------------------------------------------
         mandmin = STATMIN>0, 
        custody = PRESENT==1, 
  #------------------------------------Not all years had depart--------------------------------
        #upward = case_when(
           #BOOKERCD %in% c(1:4) ~ TRUE,
           #DEPART==1 ~ TRUE,
           #DEPART_A==1 ~ TRUE,
           #DEPART==8 ~ NA,
           #DEPART_A==8 ~ NA,
           #!BOOKERCD %in% c(1:4) & !is.na(BOOKERCD) ~ FALSE, #add 8 to the nots for consistency
           #DEPART!=1 & !is.na(DEPART) ~ FALSE,
           #DEPART_A!=1 & !is.na(DEPART_A) ~ FALSE,
           #TRUE ~ NA)
 
  #--------------------------------------------------------------------
 downgovt = case_when( #inconsistent with the coding of 8 as NA
           BOOKERCD %in% c(6, 7) ~ TRUE,
           !BOOKERCD %in% c(6, 7) & !is.na(BOOKERCD) ~ FALSE,
           TRUE ~ NA), 
 downcourt = case_when(
           BOOKERCD %in% c(8:11) ~ TRUE,
           !BOOKERCD %in% c(8:11) & !is.na(BOOKERCD) ~ FALSE,
           TRUE ~ NA),
         #subasst = case_when(
           #BOOKERCD==5 ~ TRUE,
           #DEPART %in% c(3, 5, 7, 9) ~ TRUE,
           #DEPART_A==2 ~ TRUE,
           #DEPART==8 ~ NA,
           #DEPART_A==8 ~ NA,
           #BOOKERCD !=5 ~ FALSE,
           #!DEPART %in% c(3, 5, 7, 9) & !is.na(DEPART) ~ FALSE,#add 8 to the nots for consistency
           #DEPART_A!=2 & !is.na(DEPART_A) ~ FALSE,
           #TRUE ~ NA),
         valve = case_when(
           SAFE %in% c(1, 2) ~ TRUE,
           SAFE == 0 ~ FALSE,
           TRUE ~ NA), 
        mitigate = MITROLHI!=0,
         aggravate = AGGROLHI!=0,
         agedummy = AGE>25,
         educ = case_when(
           EDUCATN %in% c(13:16, 23, 24, 34, 35) ~ TRUE,
           !EDUCATN %in% c(13:16, 23, 24, 34, 35) & !is.na(EDUCATN) ~ FALSE,
           TRUE ~ NA),
         citizen = NEWCIT==0,
         whitefemale = MONSEX==1 & NEWRACE==1,
         blackmale = MONSEX==0 & NEWRACE==2,
         blackfemale = MONSEX==1 & NEWRACE==2,
         hispmale = MONSEX==0 & NEWRACE==3,
         hispfemale = MONSEX==1 & NEWRACE==3,
         othermale = MONSEX==0 & NEWRACE==6,
         otherfemale = MONSEX==1 & NEWRACE==6, 
 racesex = case_when(
           MONSEX==0 & NEWRACE==1 ~ "whitemale",
           MONSEX==1 & NEWRACE==1 ~ "whitefemale",
           MONSEX==0 & NEWRACE==2 ~ "blackmale",
           MONSEX==1 & NEWRACE==2 ~ "blackfemale",
           MONSEX==0 & NEWRACE==3 ~ "hispmale",
           MONSEX==1 & NEWRACE==3 ~ "hispfemale",
           MONSEX==0 & NEWRACE==6 ~ "othermale",
           MONSEX==1 & NEWRACE==6 ~ "otherfemale",
           is.na(MONSEX) | is.na(NEWRACE) ~ NA_character_
         )) %>%  #NAs are messed up
  mutate( #-------------------------------------------------------------------Couldn't do othtype2 and immigration
    crime_type = case_when( # uses the refined model typology
    drugtraff ~ "Drug Trafficking",
    whitecoll ~ "White Collar",
    #othtype2 ~ "Other Types",
    sexual2 ~ "Sexual",
    porn ~ "Porn",
    violent ~ "Violent",
    #immigration ~ "Immigration"
    ),
    racesex_clean = case_when(
      str_detect(racesex, "whitemale") ~ "White Male",
      str_detect(racesex, "whitefemale") ~ "White Female",
      str_detect(racesex, "blackmale") ~ "Black Male",
      str_detect(racesex, "blackfemale") ~ "Black Female",
      str_detect(racesex, "hispmale") ~ "Hispanic Male",
      str_detect(racesex, "hispfemale") ~ "Hispanic Female",
      str_detect(racesex, "othermale") ~ "Other Male",
      str_detect(racesex, "otherfemale") ~ "Other Female"),
    #downdep = case_when(
      #DEPART %in% c(2, 4, 6) ~ TRUE,
      #DEPART_A %in% c(3, 4, 5) ~ TRUE,
      #downgovt ~ TRUE,
      #downcourt ~ TRUE,
      #DEPART==8 ~ NA,
      #DEPART_A==8 ~ NA, #consider FALSE here?
      #!DEPART %in% c(2, 4, 6) & !is.na(DEPART) ~ FALSE,
      #!DEPART_A %in% c(3, 4, 5) & !is.na(DEPART_A) ~ FALSE,
      #!downgovt & !downcourt ~ FALSE,
      #TRUE ~ NA),
    mandmin2 = case_when(
      is.na(STATMIN) ~ NA,
      STATMIN==0 ~ FALSE,
      STATMIN>0 & (SAFE==1 | SAFE==2) ~ FALSE,
      #STATMIN>0 & subasst ~ FALSE,#-----------------------------------------couldn't use subasst bc of depart
      STATMIN>0 ~ TRUE,
      TRUE ~ NA))
  #mutate(familyties = str_detect(reason, "\\b17\\b") ------------------for some reason had no reason column?
         

#write_csv(data, "~/Downloads/research/io_2017_2021.csv")

```

# Step 3: The Model 

```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               #immigration +
               #othtype2 +
               whitecoll +
               #upward +
               #downdep +
               #subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```


## 2017-2021

```{r}
model2_2017_2021 <- model2(data)

as_flextable(model2_2017_2021)
```



