---
title: "Demographic Baseline 2017-2021"
fontsize: 12pt
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    toc_depth: 5
    number_sections: true
  pdf_document: default
urlcolor: magenta
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```   

1/25/23 remaining tasks
* combine code with read_sentencing and baseline switch
* check that we're not losing any data when the new fys are read in (correct total number of rows, no systematically missing vars, etc)
* check that after data processing, we have expected numbers of NAs across vars
* re-do other reports to integrate new fys

```{r , message=FALSE}
library(bigrquery)
library(DBI)
library(tidyverse)
library(flextable)
library(janitor)
library(scales)
library(glue)
library(lubridate)
```


# Step 1: Making another io_raw.csv 


```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "balmy-coral-330818",
  dataset = "bqtest"
  )

dbListTables(con)

```


Aggregate Reasons function
```{r}
# CML: especially for functions like this, that you'll use repeatedly within a project like ours, it's important to only write out the function once so that you don't have to make edits to each place it appears! so let's source in from a utils script rather than writing it out here. 

source(here::here("utils.R"))



```

## 2017 
In fy17_1: 
SENSPLT0, GLMIN, TOTCHPTS, IS924C, WEAPSOC, STATMIN, CAROFFAP, ACCAP, SAFE, NEWCNVTN, PRESENT, MITROLHI, AGGROLHI, NEWRACE, MONSEX, AGE, EDUCATN, NEWCIT

In fy17_2: 
GDLINEHI

```{r}
query2017_1 = dbSendQuery(con, "select `id`,`SENTMON`,`SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `BOOKERCD`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`from fy17_1")
data2017_1 <-  dbFetch(query2017_1) %>% 
  aggregate_reasons()

query2017_2 <- dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy17_2 ")
data2017_2 <-  dbFetch(query2017_2)

data2017 <- full_join(data2017_1, data2017_2, by = 'id') %>% 
  mutate(opafy = 2017,
         SENTRNGE = NA) %>% 
  select(-1)


#write_csv(data2017, here::here("data/io_truncated/data2017.csv"))

```

## 2018
In opafy18nid_1: 
SENSPLT0, GLMIN, TOTCHPTS, IS924C, WEAPSOC, STATMIN, CAROFFAP, ACCAP, SAFE, NEWCNVTN, PRESENT, MITROLHI, AGGROLHI, NEWRACE, MONSEX, AGE, EDUCATN, NEWCIT, SENTRNGE

In opafy18nid_3: 
GDLINEHI
```{r}
query2018_1 <- dbSendQuery(con, "select `id`,`sentmon`, `sentyr`,`sensplt0`, `glmin`,`totchpts`,`is924c`,`weapsoc`, `statmin`, `caroffap`,`accap`, `safe`, `newcnvtn`, `present`, `mitrolhi`, `aggrolhi`, `newrace`, `monsex`, `age`, `educatn`, `newcit`, `sentrnge`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26`,`REAS27`,`REAS28`,`REAS29`,`REAS30`,`REAS31`,`REAS32`,`REAS33`,`REAS34`,`REAS35` from fy18_1")

data2018_1 <-  dbFetch(query2018_1) %>% aggregate_reasons()

query2018_2 <-  dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy18_3")
data2018_2 <- dbFetch(query2018_2)

data2018 <- full_join(data2018_1, data2018_2, by = 'id') %>% 
  mutate(opafy=2018) %>% 
  select(-1) %>% 
  rename_at(vars(sentmon:sentrnge), str_to_upper)

head(data2018)

#write_csv(data2018, here::here("data/io_truncated/data2018.csv"))
```

## 2019 
```{r}
query2019_1 <-  dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `SENTRNGE`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26`,`REAS27`,`REAS28` from fy19_1")
data2019_1 <-  dbFetch(query2019_1) %>% aggregate_reasons()
query2019_2 <-  dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy19_2 ")
data2019_2 <- dbFetch(query2019_2)

data2019 <- full_join(data2019_1, data2019_2, by = 'id') %>% 
  mutate(opafy=2019) %>% 
  select(-1)

head(data2019)

#write_csv(data2019, here::here("data/io_truncated/data2019.csv"))
```

## 2020 
```{r}
query2020_1 <-  dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `SENTRNGE`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26` from fy20_1")
data2020_1 <-  dbFetch(query2020_1) %>% aggregate_reasons()

query2020_2 <-  dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy20_2 ")
data2020_2 <- dbFetch(query2020_2)

data2020 <- full_join(data2020_1, data2020_2, by = 'id') %>% 
  mutate(opafy = 2020) %>% 
  select(-1)

head(data2020)

#write_csv(data2020, here::here("data/io_truncated/data2020.csv"))
```


## 2021 
```{r}
query2021_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `SENTRNGE`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26`,`REAS27`,`REAS28`,`REAS29`,`REAS30` from fy21_1")

data2021_1 <-  dbFetch(query2021_1) %>% aggregate_reasons()
query2021_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy21_3")
data2021_2 <- dbFetch(query2021_2)

data2021 <- full_join(data2021_1, data2021_2, by = 'id') %>% 
  mutate(opafy=2021) %>% 
  select(-1)

head(data2021)

#write_csv(data2021, here::here("data/io_truncated/data2021.csv"))
```

## Merge together 
```{r, message=FALSE, warning=FALSE}
#CML rewrite as mutate

io_raw_2017_2021 <- bind_rows(data2017, data2018, data2019, data2020, data2021) %>% 
  mutate(across(c(SENTMON, SENTYR, SENSPLT0, GLMIN, TOTCHPTS, IS924C,WEAPSOC,STATMIN, CAROFFAP, ACCAP, SAFE,NEWCNVTN, PRESENT,MITROLHI,AGGROLHI,NEWRACE,MONSEX,AGE, EDUCATN,NEWCIT,BOOKERCD,SENTRNGE), as.numeric))

#write_csv(io_raw_2017_2021,here::here("data/io_truncated/io_raw_2017_2021.csv"))
```

```{r}
str(io_raw_2017_2021)
```


**Depart Variable**
In the USSC Public Release Codebook, the variable for Departure states: 
- BOOKERCD is the variable generally used for outside the range analysis post-Booker
- 2004-2017 uses either DEPART_A, BOOKERCD, or DEPART
- 2018-PRESENT uses SENTRNGE 
- For some reason, there is no column like DEPART in 2017. DSPSR?   


# Step 2: Creating io 2017-2021 

```{r}

##CML: I haven't looked yet -- let me know when you get to this point
matchwna <- function(string, list){
  case_when(
    string %in% list ~ TRUE,
    !is.na(string) ~ FALSE,
    is.na(string) ~ NA
  )
}


data <- io_raw_2017_2021 %>% 
  mutate(SENTMON = SENTMON %>% 
           str_replace("01", "Jan") %>% 
           str_replace("02","Feb" ) %>% 
           str_replace("03","Mar") %>% 
           str_replace("04","Apr") %>% 
           str_replace("05","May") %>% 
           str_replace("06","Jun") %>% 
           str_replace("07","Jul") %>% 
           str_replace("08","Aug") %>% 
           str_replace("09","Sep") %>% 
           str_replace("10","Oct") %>% 
           str_replace("11","Nov") %>% 
           str_replace("12","Dec"), 
          sentmonyr = ymd(glue("{SENTYR}-{SENTMON}-01")), #5 failed to parse 
         #CML: why convert the numeric months to "Jan", "Feb" etc? it's also only working for some rows, because the data seems to be encoded as "1" instead of "01" for instance
         logsplit = case_when(
          SENSPLT0==0.00 ~ log(0.01), 
          SENSPLT0>470 ~ log(470),
          TRUE ~ log(SENSPLT0)),
         logmin = case_when(
          GLMIN==0.00 ~ log(0.01), 
          GLMIN>470 ~ log(470),
          TRUE ~ log(GLMIN)), 
         
         violent = matchwna(GDLINEHI, c("2A1.1", "2A1.2", "2A1.3", "2A1.4", "2A1.5", "2A2.1", "2A2.2", "2A2.3", "2A2.4", 
                                        "2A4.1", "2A4.2", "2A5.1", "2A5.2", "2A5.3", "2A6.1", "2A6.2", "2E1.3", "2E1.4", 
                                        "2E2.1", "2B3.1", "2B3.2", "2B3.3")) | str_detect(GDLINEHI, "^2K\\d{1}\\.\\d{1}$"), 
         sexual = matchwna(GDLINEHI, c("2A3.1", "2A3.2", "2A3.3", "2A3.4", "2G1.1", "2G1.2", "2G1.3", "2G2.1", "2G2.2", "2G2.3", "2G2.4", "2G2.5")), 
         sexual2 = matchwna(GDLINEHI, c("2A3.1", "2A3.2", "2A3.3", "2A3.4", "2G1.1", "2G1.2", "2G1.3")),
         porn = matchwna(GDLINEHI, c("2G2.1", "2G2.2", "2G2.3", "2G2.4", "2G2.5")),
         drugtraff = matchwna(GDLINEHI, c("2D1.1", "2D1.2", "2D1.5", "2D1.6", "2D1.7", "2D1.8", "2D1.9", "2D1.10", "2D1.11", "2D1.12", "2D1.13")),
         othdrug = matchwna(GDLINEHI, c("2D2.1", "2D2.2")),
         whitecoll = matchwna(GDLINEHI, c("2B1.1", "2B1.6", "2B4.1", "2B5.1", "2B5.3", "2F1.1", "2F1.2", "2R1.1", "2S1.1", "2S1.2", "2S1.3", "2S1.4")) | 
           str_detect(GDLINEHI, "^2T\\d{1}\\.\\d{1}$"), 
         immigration = str_detect(GDLINEHI, "^2L\\d{1}\\.\\d{1}$"), # one of the main changes! #CML: what is? immigration var looks the same. just removing the time periods we had in the old one? also note to include a new time period for this data when we integrate with the main data file
      drug = drugtraff | othdrug,
         othtype = !(drugtraff|othdrug|violent|sexual|whitecoll|immigration), 
         othtype2 = !(violent|sexual2|porn|drugtraff|whitecoll|immigration),  
         mandmin = STATMIN>0,
         custody = PRESENT==1,
         upward = case_when( # another big change here! 
           BOOKERCD %in% c(1:4) ~ TRUE,
           SENTRNGE %in% c(1, 6) ~ TRUE,
           !BOOKERCD %in% c(1:4) & !is.na(BOOKERCD) ~ FALSE, 
           !SENTRNGE %in% c(1, 6) & !is.na(SENTRNGE) ~ FALSE,
           TRUE ~ NA),
         downgovt = case_when( 
           BOOKERCD %in% c(6, 7) ~ TRUE,
           SENTRNGE %in% c(3, 4, 7) ~ TRUE,
           !BOOKERCD %in% c(6, 7) & !is.na(BOOKERCD) ~ FALSE,
           !SENTRNGE %in% c(3, 4, 7) & !is.na(SENTRNGE) ~ FALSE, #CML: double check that early disposition/5K3.1 should be classified as downgovt and not downcourt
           TRUE ~ NA), 
         downcourt = case_when(
           BOOKERCD %in% c(8:11) ~ TRUE,
           SENTRNGE %in% c(5, 8) ~ TRUE,
           !BOOKERCD %in% c(8:11) & !is.na(BOOKERCD) ~ FALSE,
           !SENTRNGE %in% c(5, 8) & !is.na(SENTRNGE) ~ FALSE,
           TRUE ~ NA),
         subasst = case_when(
           BOOKERCD==5 ~ TRUE,
           SENTRNGE ==2 ~ TRUE,
           BOOKERCD !=5 ~ FALSE,
           SENTRNGE !=2 ~ FALSE,
           TRUE ~ NA), 
         valve = case_when( #CML: may not fully capture valve cases post-2019, see codebook
           SAFE %in% c(1, 2) ~ TRUE,
           SAFE == 0 ~ FALSE,
           TRUE ~ NA),
         mitigate = MITROLHI!=0,
         aggravate = AGGROLHI!=0,
         agedummy = AGE>25,
         educ = case_when(
           EDUCATN %in% c(13:16, 23, 24, 34, 35) ~ TRUE,
           !EDUCATN %in% c(13:16, 23, 24, 34, 35) & !is.na(EDUCATN) ~ FALSE,
           TRUE ~ NA),
         citizen = NEWCIT==0,
         whitefemale = MONSEX==1 & NEWRACE==1,
         blackmale = MONSEX==0 & NEWRACE==2,
         blackfemale = MONSEX==1 & NEWRACE==2,
         hispmale = MONSEX==0 & NEWRACE==3,
         hispfemale = MONSEX==1 & NEWRACE==3,
         othermale = MONSEX==0 & NEWRACE==6,
         otherfemale = MONSEX==1 & NEWRACE==6, 
 racesex = case_when(
           MONSEX==0 & NEWRACE==1 ~ "whitemale",
           MONSEX==1 & NEWRACE==1 ~ "whitefemale",
           MONSEX==0 & NEWRACE==2 ~ "blackmale",
           MONSEX==1 & NEWRACE==2 ~ "blackfemale",
           MONSEX==0 & NEWRACE==3 ~ "hispmale",
           MONSEX==1 & NEWRACE==3 ~ "hispfemale",
           MONSEX==0 & NEWRACE==6 ~ "othermale",
           MONSEX==1 & NEWRACE==6 ~ "otherfemale",
           is.na(MONSEX) | is.na(NEWRACE) ~ NA_character_
         )) %>%  
  mutate( 
    crime_type = case_when( 
    drugtraff ~ "Drug Trafficking",
    whitecoll ~ "White Collar",
    othtype2 ~ "Other Types",
    sexual2 ~ "Sexual",
    porn ~ "Porn",
    violent ~ "Violent",
    immigration ~ "Immigration"
    ),
    racesex_clean = case_when(
      str_detect(racesex, "whitemale") ~ "White Male",
      str_detect(racesex, "whitefemale") ~ "White Female",
      str_detect(racesex, "blackmale") ~ "Black Male",
      str_detect(racesex, "blackfemale") ~ "Black Female",
      str_detect(racesex, "hispmale") ~ "Hispanic Male",
      str_detect(racesex, "hispfemale") ~ "Hispanic Female",
      str_detect(racesex, "othermale") ~ "Other Male",
      str_detect(racesex, "otherfemale") ~ "Other Female"),
    downdep = case_when( ###this is called downdep but really it's all below range sentences, not just those attributable only to departures, so in addition to 3,4,5 you'd want  7, 8 (which is not covered by downcourt and downgovt)
      downgovt ~ TRUE,
      downcourt ~ TRUE,
      !downgovt & !downcourt ~ FALSE,
      TRUE ~ NA),
    mandmin2 = case_when(
      is.na(STATMIN) ~ NA,
      STATMIN==0 ~ FALSE,
      STATMIN>0 & (SAFE==1 | SAFE==2) ~ FALSE,
      STATMIN>0 & subasst ~ FALSE,
      STATMIN>0 ~ TRUE,
      TRUE ~ NA),
  familyties =  str_detect(reason, "\\b17\\b"))
         

write_csv(data, here::here("data/io_2017_2021.csv"))

```

# Step 3: The Model 

```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```


## 2017-2021

```{r}
model2_2017_2021 <- model2(data)

as_flextable(model2_2017_2021)
```

# Plots 
## Functions 
```{r}
demogbar <- function(df){
  df %>% 
    ggplot(aes(x = period, y = value, fill = period))+
    geom_bar(stat="identity")+
    geom_text(aes(label=value), vjust=.5, size = 6)+ 
    labs(y = "Percent", x = "Period",
         caption = "Values not statistically significant at the 0.01 level are omitted.")+
    theme_minimal() + 
    facet_wrap(~demog)+
    scale_fill_brewer(type = "qual", palette = 6) +
    theme(strip.text = element_text(size = 13),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          plot.caption = element_text(size = 11),
          axis.text.y = element_blank(),
          axis.text.x=element_blank())
  
} 
extract_coeffs_demog <- function(df){
  df %>%
    extract_coeffs() %>% 
    filter(str_detect(var, "male")) %>% 
    mutate(value = logtrans(coeff),
           demog = case_when(
            str_detect(var, "whitemale") ~ "White Male",
            str_detect(var, "whitefemale") ~ "White Female",
            str_detect(var, "blackmale") ~ "Black Male",
            str_detect(var, "blackfemale") ~ "Black Female",
            str_detect(var, "hispmale") ~ "Hispanic Male",
            str_detect(var, "hispfemale") ~ "Hispanic Female",
            str_detect(var, "othermale") ~ "Other Male",
            str_detect(var, "otherfemale") ~ "Other Female")) %>% 
    select(demog, value)
}
```

## White Male Baseline 
Because the original model uses white male as the baseline, we use model2_2017_2021 here 
```{r}

present_wm <- extract_coeffs_demog(model2_2017_2021) %>% mutate(period = "Present")
present_wm %>% filter(str_detect(demog, "Female")) %>% demogbar()

present_wm %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

##  White Female Baseline

```{r}
model2_wf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 7) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_present_wf <- model2_wf(data)
present_wf <- extract_coeffs_demog(model2_present_wf) %>% mutate(period = "Present")

present_wf %>% filter(str_detect(demog, "Female")) %>% demogbar()
present_wf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

## Hispanic Female Baseline

```{r}
model2_hf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 3) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_present_hf <- model2_hf(data)
present_hf <- extract_coeffs_demog(model2_present_hf) %>% mutate(period = "Present")
present_hf %>% filter(str_detect(demog, "Female")) %>% demogbar()
present_hf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

## Black Female Baseline

```{r}
model2_bf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 1) +
               agedummy +
               educ +
               citizen,
             df)
}
```
  
```{r}
model2_present_bf <- model2_bf(data)
present_bf <- extract_coeffs_demog(model2_present_bf) %>% mutate(period = "Present")
present_bf%>% filter(str_detect(demog, "Female")) %>% demogbar()
present_bf%>% filter(str_detect(demog, "Male")) %>% demogbar()
```




