---
title: "Demographic Baseline 2017-2021"
fontsize: 12pt
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    toc_depth: 5
    number_sections: true
  pdf_document: default
urlcolor: magenta
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```   

1/25/23 remaining tasks
* combine code with read_sentencing and baseline switch
  * make new .R file with all of the bigquery setup that you have at the top here (so, everything through the data <- io_raw_2017_2021 step) 02_io_download
  * don't forget to source in the utils file and run the aggregate_reasons function over each dataset before you merge them (source(here::here("utils.R")))
  * have the output of that file save to here::here(data/io_raw_2017_2021.csv)
  * integrate changes made in your data processing script (vis a vis original read_sentencing)
* check that we're not losing any data when the new fys are read in (correct total number of rows, no systematically missing vars, etc)
* check that after data processing, we have expected numbers of NAs across vars
* re-do other reports to integrate new fys

```{r , message=FALSE}
library(tidyverse)
library(flextable)
library(janitor)
library(scales)
library(glue)
library(lubridate)
source(here::here("utils.R"))
```

# The Model 

```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```


## 2017-2021

```{r}
model2_2017_2021 <- model2(data)

as_flextable(model2_2017_2021)
```

# Plots 
## Functions 
```{r}
demogbar <- function(df){
  df %>% 
    ggplot(aes(x = period, y = value, fill = period))+
    geom_bar(stat="identity")+
    geom_text(aes(label=value), vjust=.5, size = 6)+ 
    labs(y = "Percent", x = "Period",
         caption = "Values not statistically significant at the 0.01 level are omitted.")+
    theme_minimal() + 
    facet_wrap(~demog)+
    scale_fill_brewer(type = "qual", palette = 6) +
    theme(strip.text = element_text(size = 13),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          plot.caption = element_text(size = 11),
          axis.text.y = element_blank(),
          axis.text.x=element_blank())
  
} 
extract_coeffs_demog <- function(df){
  df %>%
    extract_coeffs() %>% 
    filter(str_detect(var, "male")) %>% 
    mutate(value = logtrans(coeff),
           demog = case_when(
            str_detect(var, "whitemale") ~ "White Male",
            str_detect(var, "whitefemale") ~ "White Female",
            str_detect(var, "blackmale") ~ "Black Male",
            str_detect(var, "blackfemale") ~ "Black Female",
            str_detect(var, "hispmale") ~ "Hispanic Male",
            str_detect(var, "hispfemale") ~ "Hispanic Female",
            str_detect(var, "othermale") ~ "Other Male",
            str_detect(var, "otherfemale") ~ "Other Female")) %>% 
    select(demog, value)
}
```

## White Male Baseline 
Because the original model uses white male as the baseline, we use model2_2017_2021 here 
```{r}

present_wm <- extract_coeffs_demog(model2_2017_2021) %>% mutate(period = "Present")
present_wm %>% filter(str_detect(demog, "Female")) %>% demogbar()

present_wm %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

##  White Female Baseline

```{r}
model2_wf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 7) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_present_wf <- model2_wf(data)
present_wf <- extract_coeffs_demog(model2_present_wf) %>% mutate(period = "Present")

present_wf %>% filter(str_detect(demog, "Female")) %>% demogbar()
present_wf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

## Hispanic Female Baseline

```{r}
model2_hf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 3) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_present_hf <- model2_hf(data)
present_hf <- extract_coeffs_demog(model2_present_hf) %>% mutate(period = "Present")
present_hf %>% filter(str_detect(demog, "Female")) %>% demogbar()
present_hf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

## Black Female Baseline

```{r}
model2_bf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 1) +
               agedummy +
               educ +
               citizen,
             df)
}
```
  
```{r}
model2_present_bf <- model2_bf(data)
present_bf <- extract_coeffs_demog(model2_present_bf) %>% mutate(period = "Present")
present_bf%>% filter(str_detect(demog, "Female")) %>% demogbar()
present_bf%>% filter(str_detect(demog, "Male")) %>% demogbar()
```




