---
title: "Demographic Baseline 2017-2021"
fontsize: 12pt
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    toc_depth: 5
    number_sections: true
  pdf_document: default
urlcolor: magenta
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```   


```{r , message=FALSE}
library(bigrquery)
library(DBI)
library(tidyverse)
library(flextable)
library(janitor)
library(scales)
library(glue)
library(lubridate)
```


# Step 1: Making another io_raw.csv 


```{r}
con <- dbConnect(
  bigrquery::bigquery(),
  project = "balmy-coral-330818",
  dataset = "bqtest"
  )

dbListTables(con)

```


Aggregate Reasons function
```{r}
# CML: especially for functions like this, that you'll use repeatedly within a project like ours, it's important to only write out the function once so that you don't have to make edits to each place it appears! so let's source in from a utils script rather than writing it out here. 

source(here::here("utils.R"))



```

## 2017 
In fy17_1: 
SENSPLT0, GLMIN, TOTCHPTS, IS924C, WEAPSOC, STATMIN, CAROFFAP, ACCAP, SAFE, NEWCNVTN, PRESENT, MITROLHI, AGGROLHI, NEWRACE, MONSEX, AGE, EDUCATN, NEWCIT

In fy17_2: 
GDLINEHI

```{r}
query2017_1 = dbSendQuery(con, "select `id`,`SENTMON`,`SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `BOOKERCD`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`from fy17_1")
data2017_1 <-  dbFetch(query2017_1)
data2017_1 <- aggregate_reasons(data2017_1) 
head(data2017_1)

query2017_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy17_2 ")
data2017_2 = dbFetch(query2017_2)
head(data2017_2)

data2017 <- merge(data2017_1, data2017_2, by = 'id')
data2017$opafy <- 2017

data2017 <- subset(data2017, select= -c(1))

data2017$SENTRNGE <- NA
head(data2017)

#write.csv(data2017, "~/Downloads/doherty_research/data_files/data2017.csv")

```

## 2018
In opafy18nid_1: 
SENSPLT0, GLMIN, TOTCHPTS, IS924C, WEAPSOC, STATMIN, CAROFFAP, ACCAP, SAFE, NEWCNVTN, PRESENT, MITROLHI, AGGROLHI, NEWRACE, MONSEX, AGE, EDUCATN, NEWCIT, SENTRNGE

In opafy18nid_3: 
GDLINEHI
```{r}
query2018_1 = dbSendQuery(con, "select `id`,`sentmon`, `sentyr`,`sensplt0`, `glmin`,`totchpts`,`is924c`,`weapsoc`, `statmin`, `caroffap`,`accap`, `safe`, `newcnvtn`, `present`, `mitrolhi`, `aggrolhi`, `newrace`, `monsex`, `age`, `educatn`, `newcit`, `sentrnge`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26`,`REAS27`,`REAS28`,`REAS29`,`REAS30`,`REAS31`,`REAS32`,`REAS33`,`REAS34`,`REAS35` from fy18_1")
data2018_1 <-  dbFetch(query2018_1)
data2018_1 <- aggregate_reasons(data2018_1,35)
head(data2018_1)

query2018_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy18_3")
data2018_2 <- dbFetch(query2018_2)
head(data2018_2)

data2018 <- merge(data2018_1, data2018_2, by = 'id')

names(data2018) <- toupper(names(data2018))
names(data2018)[23] <- tolower(names(data2018)[23])

data2018$opafy <- 2018
data2018 <- subset(data2018, select= -c(1))
head(data2018)

#write.csv(data2018, "~/Downloads/doherty_research/data_files/data2018.csv")
```

## 2019 
```{r}
query2019_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `SENTRNGE`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26`,`REAS27`,`REAS28` from fy19_1")
data2019_1 <-  dbFetch(query2019_1)
data2019_1 <- aggregate_reasons(data2019_1, 28)
head(data2019_1)

query2019_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy19_2 ")
data2019_2 <- dbFetch(query2019_2)
head(data2019_2)

data2019 <- merge(data2019_1, data2019_2, by = 'id')
data2019$opafy <- 2019
data2019 <- subset(data2019, select= -c(1))
head(data2019)

#write.csv(data2019, "~/Downloads/doherty_research/data_files/data2019.csv")
```

## 2020 
```{r}
query2020_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `SENTRNGE`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26` from fy20_1")
data2020_1 <-  dbFetch(query2020_1)
data2020_1 <- aggregate_reasons(data2020_1, 26)
head(data2020_1)

query2020_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy20_2 ")
data2020_2 <- dbFetch(query2020_2)
head(data2020_2)

data2020 <- merge(data2020_1, data2020_2, by = 'id')
data2020$opafy <- 2020
data2020 <- subset(data2020, select= -c(1))
head(data2020)

#write.csv(data2020, "~/Downloads/doherty_research/data_files/data2020.csv")
```


## 2021 
```{r}
query2021_1 = dbSendQuery(con, "select `id`,`SENTMON`, `SENTYR`,`SENSPLT0`, `GLMIN`,`TOTCHPTS`,`IS924C`,`WEAPSOC`, `STATMIN`, `CAROFFAP`,`ACCAP`, `SAFE`, `NEWCNVTN`, `PRESENT`, `MITROLHI`, `AGGROLHI`, `NEWRACE`, `MONSEX`, `AGE`, `EDUCATN`, `NEWCIT`, `SENTRNGE`,`REAS1`, `REAS2`, `REAS3`, `REAS4`, `REAS5`, `REAS6`, `REAS7`, `REAS8`, `REAS9`, `REAS10`, `REAS11`, `REAS12`, `REAS13`, `REAS14`, `REAS15`, `REAS16`, `REAS17`,`REAS18`,`REAS19`,`REAS20`,`REAS21`, `REAS22`, `REAS23`, `REAS24`, `REAS25`,`REAS26`,`REAS27`,`REAS28`,`REAS29`,`REAS30` from fy21_1")

data2021_1 <-  dbFetch(query2021_1)
data2021_1 <- aggregate_reasons(data2021_1, 30)
head(data2021_1)

query2021_2 = dbSendQuery(con, "select `id`, `GDLINEHI` FROM fy21_3")
data2021_2 <- dbFetch(query2021_2)
head(data2021_2)

data2021 <- merge(data2021_1, data2021_2, by = 'id')
data2021$opafy <- 2021
data2021 <- subset(data2021, select= -c(1))
head(data2021)

#write.csv(data2021, "~/Downloads/doherty_research/data_files/data2021.csv")
```

## Merge together 
```{r, message=FALSE, warning=FALSE}
io_raw_2017_2021 <- bind_rows(data2017, data2018, data2019, data2020, data2021)

io_raw_2017_2021$SENTMON <- as.numeric(io_raw_2017_2021$SENTMON)
io_raw_2017_2021$SENTYR <- as.numeric(io_raw_2017_2021$SENTYR)
io_raw_2017_2021$SENSPLT0 <- as.numeric(io_raw_2017_2021$SENSPLT0)
io_raw_2017_2021$GLMIN <-as.numeric( io_raw_2017_2021$GLMIN)
io_raw_2017_2021$TOTCHPTS <- as.numeric(io_raw_2017_2021$TOTCHPTS)
io_raw_2017_2021$IS924C <- as.numeric(io_raw_2017_2021$IS924C)
io_raw_2017_2021$WEAPSOC <- as.numeric(io_raw_2017_2021$WEAPSOC)
io_raw_2017_2021$STATMIN <- as.numeric(io_raw_2017_2021$STATMIN)
io_raw_2017_2021$CAROFFAP <- as.numeric(io_raw_2017_2021$CAROFFAP)
io_raw_2017_2021$ACCAP <- as.numeric(io_raw_2017_2021$ACCAP)
io_raw_2017_2021$SAFE <- as.numeric(io_raw_2017_2021$SAFE)#
io_raw_2017_2021$NEWCNVTN <- as.numeric(io_raw_2017_2021$NEWCNVTN)
io_raw_2017_2021$PRESENT <- as.numeric(io_raw_2017_2021$PRESENT)
io_raw_2017_2021$MITROLHI <- as.numeric(io_raw_2017_2021$MITROLHI)
io_raw_2017_2021$AGGROLHI <- as.numeric(io_raw_2017_2021$AGGROLHI)
io_raw_2017_2021$NEWRACE <- as.numeric(io_raw_2017_2021$NEWRACE)
io_raw_2017_2021$MONSEX <- as.numeric(io_raw_2017_2021$MONSEX)
io_raw_2017_2021$AGE <- as.numeric(io_raw_2017_2021$AGE)
io_raw_2017_2021$EDUCATN <- as.numeric(io_raw_2017_2021$EDUCATN)
io_raw_2017_2021$NEWCIT <- as.numeric(io_raw_2017_2021$NEWCIT)
io_raw_2017_2021$BOOKERCD <- as.numeric(io_raw_2017_2021$BOOKERCD)
io_raw_2017_2021$SENTRNGE <- as.numeric(io_raw_2017_2021$SENTRNGE)

#write.csv(io_raw_2017_2021,"~/Downloads/doherty_research/data_files/io_raw_2017_2021.csv")
```

```{r}

str(io_raw_2017_2021)
```


**Depart Variable**
In the USSC Public Release Codebook, the variable for Departure states: 
- BOOKERCD is the variable generally used for outside the range analysis post-Booker
- 2004-2017 uses either DEPART_A, BOOKERCD, or DEPART
- 2018-PRESENT uses SENTRNGE 
- For some reason, there is no column like DEPART in 2017. DSPSR?   


# Step 2: Creating io 2017-2021 

```{r}
matchwna <- function(string, list){
  case_when(
    string %in% list ~ TRUE,
    !is.na(string) ~ FALSE,
    is.na(string) ~ NA
  )
}


data <- io_raw_2017_2021 %>% 
  mutate(SENTMON = SENTMON %>% 
           str_replace("01", "Jan") %>% 
           str_replace("02","Feb" ) %>% 
           str_replace("03","Mar") %>% 
           str_replace("04","Apr") %>% 
           str_replace("05","May") %>% 
           str_replace("06","Jun") %>% 
           str_replace("07","Jul") %>% 
           str_replace("08","Aug") %>% 
           str_replace("09","Sep") %>% 
           str_replace("10","Oct") %>% 
           str_replace("11","Nov") %>% 
           str_replace("12","Dec"), 
          sentmonyr = ymd(glue("{SENTYR}-{SENTMON}-01")), #5 failed to parse 
         logsplit = case_when(
          SENSPLT0==0.00 ~ log(0.01), 
          SENSPLT0>470 ~ log(470),
          TRUE ~ log(SENSPLT0)),
         logmin = case_when(
          GLMIN==0.00 ~ log(0.01), 
          GLMIN>470 ~ log(470),
          TRUE ~ log(GLMIN)), 
         
         violent = matchwna(GDLINEHI, c("2A1.1", "2A1.2", "2A1.3", "2A1.4", "2A1.5", "2A2.1", "2A2.2", "2A2.3", "2A2.4", 
                                        "2A4.1", "2A4.2", "2A5.1", "2A5.2", "2A5.3", "2A6.1", "2A6.2", "2E1.3", "2E1.4", 
                                        "2E2.1", "2B3.1", "2B3.2", "2B3.3")) | str_detect(GDLINEHI, "^2K\\d{1}\\.\\d{1}$"), 
         sexual = matchwna(GDLINEHI, c("2A3.1", "2A3.2", "2A3.3", "2A3.4", "2G1.1", "2G1.2", "2G1.3", "2G2.1", "2G2.2", "2G2.3", "2G2.4", "2G2.5")), 
         sexual2 = matchwna(GDLINEHI, c("2A3.1", "2A3.2", "2A3.3", "2A3.4", "2G1.1", "2G1.2", "2G1.3")),
         porn = matchwna(GDLINEHI, c("2G2.1", "2G2.2", "2G2.3", "2G2.4", "2G2.5")),
         drugtraff = matchwna(GDLINEHI, c("2D1.1", "2D1.2", "2D1.5", "2D1.6", "2D1.7", "2D1.8", "2D1.9", "2D1.10", "2D1.11", "2D1.12", "2D1.13")),
         othdrug = matchwna(GDLINEHI, c("2D2.1", "2D2.2")),
         whitecoll = matchwna(GDLINEHI, c("2B1.1", "2B1.6", "2B4.1", "2B5.1", "2B5.3", "2F1.1", "2F1.2", "2R1.1", "2S1.1", "2S1.2", "2S1.3", "2S1.4")) | 
           str_detect(GDLINEHI, "^2T\\d{1}\\.\\d{1}$"), 
         immigration = str_detect(GDLINEHI, "^2L\\d{1}\\.\\d{1}$"), # one of the main changes! 
      drug = drugtraff | othdrug,
         othtype = !(drugtraff|othdrug|violent|sexual|whitecoll|immigration), 
         othtype2 = !(violent|sexual2|porn|drugtraff|whitecoll|immigration),  
         mandmin = STATMIN>0,
         custody = PRESENT==1,
         upward = case_when( # another big change here! 
           BOOKERCD %in% c(1:4) ~ TRUE,
           SENTRNGE==1 ~ TRUE,
           !BOOKERCD %in% c(1:4) & !is.na(BOOKERCD) ~ FALSE, 
           SENTRNGE!=1 & !is.na(SENTRNGE) ~ FALSE,
           TRUE ~ NA),
 downgovt = case_when( 
           BOOKERCD %in% c(6, 7) ~ TRUE,
           !BOOKERCD %in% c(6, 7) & !is.na(BOOKERCD) ~ FALSE,
           TRUE ~ NA), 
 downcourt = case_when(
           BOOKERCD %in% c(8:11) ~ TRUE,
           !BOOKERCD %in% c(8:11) & !is.na(BOOKERCD) ~ FALSE,
           TRUE ~ NA),
 subasst = case_when(
           BOOKERCD==5 ~ TRUE,
           SENTRNGE ==2 ~ TRUE,
           BOOKERCD !=5 ~ FALSE,
           !SENTRNGE ==2 & !is.na(SENTRNGE) ~ FALSE,
           TRUE ~ NA), 
         valve = case_when(
           SAFE %in% c(1, 2) ~ TRUE,
           SAFE == 0 ~ FALSE,
           TRUE ~ NA),
         mitigate = MITROLHI!=0,
         aggravate = AGGROLHI!=0,
         agedummy = AGE>25,
         educ = case_when(
           EDUCATN %in% c(13:16, 23, 24, 34, 35) ~ TRUE,
           !EDUCATN %in% c(13:16, 23, 24, 34, 35) & !is.na(EDUCATN) ~ FALSE,
           TRUE ~ NA),
         citizen = NEWCIT==0,
         whitefemale = MONSEX==1 & NEWRACE==1,
         blackmale = MONSEX==0 & NEWRACE==2,
         blackfemale = MONSEX==1 & NEWRACE==2,
         hispmale = MONSEX==0 & NEWRACE==3,
         hispfemale = MONSEX==1 & NEWRACE==3,
         othermale = MONSEX==0 & NEWRACE==6,
         otherfemale = MONSEX==1 & NEWRACE==6, 
 racesex = case_when(
           MONSEX==0 & NEWRACE==1 ~ "whitemale",
           MONSEX==1 & NEWRACE==1 ~ "whitefemale",
           MONSEX==0 & NEWRACE==2 ~ "blackmale",
           MONSEX==1 & NEWRACE==2 ~ "blackfemale",
           MONSEX==0 & NEWRACE==3 ~ "hispmale",
           MONSEX==1 & NEWRACE==3 ~ "hispfemale",
           MONSEX==0 & NEWRACE==6 ~ "othermale",
           MONSEX==1 & NEWRACE==6 ~ "otherfemale",
           is.na(MONSEX) | is.na(NEWRACE) ~ NA_character_
         )) %>%  
  mutate( 
    crime_type = case_when( 
    drugtraff ~ "Drug Trafficking",
    whitecoll ~ "White Collar",
    othtype2 ~ "Other Types",
    sexual2 ~ "Sexual",
    porn ~ "Porn",
    violent ~ "Violent",
    immigration ~ "Immigration"
    ),
    racesex_clean = case_when(
      str_detect(racesex, "whitemale") ~ "White Male",
      str_detect(racesex, "whitefemale") ~ "White Female",
      str_detect(racesex, "blackmale") ~ "Black Male",
      str_detect(racesex, "blackfemale") ~ "Black Female",
      str_detect(racesex, "hispmale") ~ "Hispanic Male",
      str_detect(racesex, "hispfemale") ~ "Hispanic Female",
      str_detect(racesex, "othermale") ~ "Other Male",
      str_detect(racesex, "otherfemale") ~ "Other Female"),
    downdep = case_when(
      SENTRNGE %in% c(3, 4, 5) ~ TRUE,
      downgovt ~ TRUE,
      downcourt ~ TRUE,
      !SENTRNGE %in% c(3, 4, 5) & !is.na(SENTRNGE) ~ FALSE,
      !downgovt & !downcourt ~ FALSE,
      TRUE ~ NA),
    mandmin2 = case_when(
      is.na(STATMIN) ~ NA,
      STATMIN==0 ~ FALSE,
      STATMIN>0 & (SAFE==1 | SAFE==2) ~ FALSE,
      STATMIN>0 & subasst ~ FALSE,
      STATMIN>0 ~ TRUE,
      TRUE ~ NA),
  familyties =  str_detect(reason, "\\b17\\b"))
         

#write_csv(data, "~/Downloads/doherty_research/data_files/io_2017_2021.csv")

```

# Step 3: The Model 

```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```


## 2017-2021

```{r}
model2_2017_2021 <- model2(data)

as_flextable(model2_2017_2021)
```

# Plots 
## Functions 
```{r}
source("~/Downloads/doherty_research/code_2017_2021_lp/utils_coeffinterps.R")
demogbar <- function(df){
  df %>% 
    ggplot(aes(x = period, y = value, fill = period))+
    geom_bar(stat="identity")+
    geom_text(aes(label=value), vjust=.5, size = 6)+ 
    labs(y = "Percent", x = "Period",
         caption = "Values not statistically significant at the 0.01 level are omitted.")+
    theme_minimal() + 
    facet_wrap(~demog)+
    scale_fill_brewer(type = "qual", palette = 6) +
    theme(strip.text = element_text(size = 13),
          legend.text = element_text(size = 10),
          legend.title = element_text(size = 12),
          axis.title.y = element_text(size = 14),
          axis.title.x = element_text(size = 14),
          plot.caption = element_text(size = 11),
          axis.text.y = element_blank(),
          axis.text.x=element_blank())
  
} 
extract_coeffs_demog <- function(df){
  df %>%
    extract_coeffs() %>% 
    filter(str_detect(var, "male")) %>% 
    mutate(value = logtrans(coeff),
           demog = case_when(
            str_detect(var, "whitemale") ~ "White Male",
            str_detect(var, "whitefemale") ~ "White Female",
            str_detect(var, "blackmale") ~ "Black Male",
            str_detect(var, "blackfemale") ~ "Black Female",
            str_detect(var, "hispmale") ~ "Hispanic Male",
            str_detect(var, "hispfemale") ~ "Hispanic Female",
            str_detect(var, "othermale") ~ "Other Male",
            str_detect(var, "otherfemale") ~ "Other Female")) %>% 
    select(demog, value)
}
```

## White Male Baseline 
Because the original model uses white male as the baseline, we use model2_2017_2021 here 
```{r}

present_wm <- extract_coeffs_demog(model2_2017_2021) %>% mutate(period = "Present")
present_wm %>% filter(str_detect(demog, "Female")) %>% demogbar()

present_wm %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

##  White Female Baseline

```{r}
model2_wf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 7) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_present_wf <- model2_wf(data)
present_wf <- extract_coeffs_demog(model2_present_wf) %>% mutate(period = "Present")

present_wf %>% filter(str_detect(demog, "Female")) %>% demogbar()
present_wf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

## Hispanic Female Baseline

```{r}
model2_hf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 3) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}
model2_present_hf <- model2_hf(data)
present_hf <- extract_coeffs_demog(model2_present_hf) %>% mutate(period = "Present")
present_hf %>% filter(str_detect(demog, "Female")) %>% demogbar()
present_hf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```

## Black Female Baseline

```{r}
model2_bf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype2 +
               whitecoll +
               upward +
               downdep +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 1) +
               agedummy +
               educ +
               citizen,
             df)
}
```
  
```{r}
model2_present_bf <- model2_bf(data)
present_bf <- extract_coeffs_demog(model2_present_bf) %>% mutate(period = "Present")
present_bf%>% filter(str_detect(demog, "Female")) %>% demogbar()
present_bf%>% filter(str_detect(demog, "Male")) %>% demogbar()
```




