---
title: "Demographic Disparities Year by Year"
output:
    word_document: 
        reference_docx: "template/template.docx"
date: "September 17, 2024"

---

This year-by-year breakdown of the Commission's "refined model" of demographic disparities in sentencing is updated using the latest changes to our replication, stemming from our conversation over email with Ryan Cotter in summer of 2023. The refined model was current as of the Commission's 2017 report, but has been superseded by a more complex model in the 2023 report.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 14, fig.height = 8)

library(tidyverse)
library(flextable)
library(janitor)
library(scales)

data <- read_csv(here::here("data/io_2002_2023.csv")) 

fy02 <- data %>% 
  filter(sentdate>="2001-10-01" & sentdate<="2002-09-30") 
fy03 <- data %>% 
  filter(sentdate>="2002-10-01" & sentdate<="2003-09-30") 
fy04 <- data %>% 
  filter(sentmonyr>="2003-10-01" & sentmonyr<="2004-09-01") 
fy05 <- data %>% 
  filter(sentmonyr>="2004-10-01" & sentmonyr<="2005-09-01") 
fy06 <- data %>% 
  filter(sentmonyr>="2005-10-01" & sentmonyr<="2006-09-01") 
fy07 <- data %>% 
  filter(sentmonyr>="2006-10-01" & sentmonyr<="2007-09-01") 
fy08 <- data %>% 
  filter(sentmonyr>="2007-10-01" & sentmonyr<="2008-09-01") 
fy09 <- data %>% 
  filter(sentmonyr>="2008-10-01" & sentmonyr<="2009-09-01") 
fy10 <- data %>% 
  filter(sentmonyr>="2009-10-01" & sentmonyr<="2010-09-01")
fy11 <- data %>% 
  filter(sentmonyr>="2010-10-01" & sentmonyr<="2011-09-01")
fy12 <- data %>% 
  filter(sentmonyr>="2011-10-01" & sentmonyr<="2012-09-01")
fy13 <- data %>% 
  filter(sentmonyr>="2012-10-01" & sentmonyr<="2013-09-01")
fy14 <- data %>% 
  filter(sentmonyr>="2013-10-01" & sentmonyr<="2014-09-01")
fy15 <- data %>% 
  filter(sentmonyr>="2014-10-01" & sentmonyr<="2015-09-01")
fy16 <- data %>% 
  filter(sentmonyr>="2015-10-01" & sentmonyr<="2016-09-01")
fy17 <- data %>% 
  filter(sentmonyr>="2016-10-01" & sentmonyr<="2017-09-01")
fy18 <- data %>% 
  filter(sentmonyr>="2017-10-01" & sentmonyr<="2018-09-01")
fy19 <- data %>% 
  filter(sentmonyr>="2018-10-01" & sentmonyr<="2019-09-01")
fy20 <- data %>% 
  filter(sentmonyr>="2019-10-01" & sentmonyr<="2020-09-01")
fy21 <- data %>% 
  filter(sentmonyr>="2020-10-01" & sentmonyr<="2021-09-01")
fy22 <- data %>% 
  filter(sentmonyr>="2021-10-01" & sentmonyr<="2022-09-01")
fy23 <- data %>% 
  filter(sentmonyr>="2022-10-01" & sentmonyr<="2023-09-01")


source(here::here("utils.R"))

```



```{r}
model2 <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype +
               whitecoll +
               upward +
               down +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 8) +
               agedummy +
               educ +
               citizen,
             df)
}
```


```{r}
model2_factor_02 <- model2(fy02)
model2_factor_03 <- model2(fy03)
model2_factor_04 <- model2(fy04)
model2_factor_05 <- model2(fy05)
model2_factor_06 <- model2(fy06)
model2_factor_07 <- model2(fy07)
model2_factor_08 <- model2(fy08)
model2_factor_09 <- model2(fy09)
model2_factor_10 <- model2(fy10)
model2_factor_11 <- model2(fy11)
model2_factor_12 <- model2(fy12)
model2_factor_13 <- model2(fy13)
model2_factor_14 <- model2(fy14)
model2_factor_15 <- model2(fy15)
model2_factor_16 <- model2(fy16)
model2_factor_17 <- model2(fy17)
model2_factor_18 <- model2(fy18)
model2_factor_19 <- model2(fy19)
model2_factor_20 <- model2(fy20)
model2_factor_21 <- model2(fy21)
model2_factor_22 <- model2(fy22)
model2_factor_23 <- model2(fy23)

```


<br>
  

```{r}

demogbar <- function(df){
  df %>% 
    ggplot(aes(x = period, y = value, fill = period))+
    geom_bar(stat="identity")+
    geom_text(aes(label=round(value)), vjust=0.3, size = 4.5)+ 
    labs(y = "Percent", x = "Period",
         caption = "Values not statistically significant at the 0.01 level are omitted.")+
    theme_minimal() + 
    facet_wrap(~demog)+
    theme(strip.text = element_text(size = 25),
          legend.text = element_text(size = 20),
          legend.title = element_text(size = 22),
          axis.title.y = element_text(size = 22),
          axis.title.x = element_text(size = 22),
          plot.caption = element_text(size = 20),
          axis.text.y = element_blank(),
          axis.text.x=element_blank())
  
}

```


### White Male Baseline

```{r}

extract_coeffs_demog <- function(df){
  df %>%
    extract_coeffs_5() %>% 
    filter(str_detect(var, "male")) %>% 
    mutate(value = logtrans(coeff),
           demog = case_when(
            str_detect(var, "whitemale") ~ "White Male",
            str_detect(var, "whitefemale") ~ "White Female",
            str_detect(var, "blackmale") ~ "Black Male",
            str_detect(var, "blackfemale") ~ "Black Female",
            str_detect(var, "hispmale") ~ "Hispanic Male",
            str_detect(var, "hispfemale") ~ "Hispanic Female",
            str_detect(var, "othermale") ~ "Other Male",
            str_detect(var, "otherfemale") ~ "Other Female")) %>% 
    select(demog, value) %>% 
    filter(!str_detect(demog, "Other"))
}

fy02_wm <- extract_coeffs_demog(model2_factor_02) %>% mutate(period = "FY02")
fy03_wm <- extract_coeffs_demog(model2_factor_03) %>% mutate(period = "FY03")
fy04_wm <- extract_coeffs_demog(model2_factor_04) %>% mutate(period = "FY04")
fy05_wm <- extract_coeffs_demog(model2_factor_05) %>% mutate(period = "FY05")
fy06_wm <- extract_coeffs_demog(model2_factor_06) %>% mutate(period = "FY06")
fy07_wm <- extract_coeffs_demog(model2_factor_07) %>% mutate(period = "FY07")
fy08_wm <- extract_coeffs_demog(model2_factor_08) %>% mutate(period = "FY08")
fy09_wm <- extract_coeffs_demog(model2_factor_09) %>% mutate(period = "FY09")
fy10_wm <- extract_coeffs_demog(model2_factor_10) %>% mutate(period = "FY10")
fy11_wm <- extract_coeffs_demog(model2_factor_11) %>% mutate(period = "FY11")
fy12_wm <- extract_coeffs_demog(model2_factor_12) %>% mutate(period = "FY12")
fy13_wm <- extract_coeffs_demog(model2_factor_13) %>% mutate(period = "FY13")
fy14_wm <- extract_coeffs_demog(model2_factor_14) %>% mutate(period = "FY14")
fy15_wm <- extract_coeffs_demog(model2_factor_15) %>% mutate(period = "FY15")
fy16_wm <- extract_coeffs_demog(model2_factor_16) %>% mutate(period = "FY16")
fy17_wm <- extract_coeffs_demog(model2_factor_17) %>% mutate(period = "FY17")
fy18_wm <- extract_coeffs_demog(model2_factor_18) %>% mutate(period = "FY18")
fy19_wm <- extract_coeffs_demog(model2_factor_19) %>% mutate(period = "FY19")
fy20_wm <- extract_coeffs_demog(model2_factor_20) %>% mutate(period = "FY20")
fy21_wm <- extract_coeffs_demog(model2_factor_21) %>% mutate(period = "FY21")
fy22_wm <- extract_coeffs_demog(model2_factor_22) %>% mutate(period = "FY22")
fy23_wm <- extract_coeffs_demog(model2_factor_23) %>% mutate(period = "FY23")


bar_data_wm <- bind_rows(fy02_wm,
                          fy03_wm, 
                          fy04_wm,
                          fy05_wm,
                          fy06_wm,
                          fy07_wm,
                          fy08_wm,
                          fy09_wm,
                          fy10_wm,
                          fy11_wm,
                          fy12_wm,
                          fy13_wm, 
                          fy14_wm,
                          fy15_wm,
                          fy16_wm,
                          fy17_wm,
                          fy18_wm,
                          fy19_wm,
                          fy20_wm,
                          fy21_wm,
                          fy22_wm,
                          fy23_wm) 
# 
# %>% 
#   mutate(period = fct_relevel(period, "Protect", "Booker", "Gall", "Post-Report", "Present"))
```

```{r}
bar_data_wm %>% filter(str_detect(demog, "Female")) %>% demogbar()
```
  
  
```{r}
bar_data_wm %>% filter(str_detect(demog, "Male")) %>% demogbar()

#add in statistical significance dots
#fix position of bar labels

```
<br>
<br>


### White Female Baseline


```{r}
model2_wf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype +
               whitecoll +
               upward +
               down +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 7) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}

model2_02_wf <- model2_wf(fy02)
model2_03_wf <- model2_wf(fy03)
model2_04_wf <- model2_wf(fy04)
model2_05_wf <- model2_wf(fy05)
model2_06_wf <- model2_wf(fy06)
model2_07_wf <- model2_wf(fy07)
model2_08_wf <- model2_wf(fy08)
model2_09_wf <- model2_wf(fy09)
model2_10_wf <- model2_wf(fy10)
model2_11_wf <- model2_wf(fy11)
model2_12_wf <- model2_wf(fy12)
model2_13_wf <- model2_wf(fy13)
model2_14_wf <- model2_wf(fy14)
model2_15_wf <- model2_wf(fy15)
model2_16_wf <- model2_wf(fy16)
model2_17_wf <- model2_wf(fy17)
model2_18_wf <- model2_wf(fy18)
model2_19_wf <- model2_wf(fy19)
model2_20_wf <- model2_wf(fy20)
model2_21_wf <- model2_wf(fy21)
model2_22_wf <- model2_wf(fy22)
model2_23_wf <- model2_wf(fy23)

```

```{r}

fy02_wf <- extract_coeffs_demog(model2_02_wf) %>% mutate(period = "FY02")
fy03_wf <- extract_coeffs_demog(model2_03_wf) %>% mutate(period = "FY03")
fy04_wf <- extract_coeffs_demog(model2_04_wf) %>% mutate(period = "FY04")
fy05_wf <- extract_coeffs_demog(model2_05_wf) %>% mutate(period = "FY05")
fy06_wf <- extract_coeffs_demog(model2_06_wf) %>% mutate(period = "FY06")
fy07_wf <- extract_coeffs_demog(model2_07_wf) %>% mutate(period = "FY07")
fy08_wf <- extract_coeffs_demog(model2_08_wf) %>% mutate(period = "FY08")
fy09_wf <- extract_coeffs_demog(model2_09_wf) %>% mutate(period = "FY09")
fy10_wf <- extract_coeffs_demog(model2_10_wf) %>% mutate(period = "FY10")
fy11_wf <- extract_coeffs_demog(model2_11_wf) %>% mutate(period = "FY11")
fy12_wf <- extract_coeffs_demog(model2_12_wf) %>% mutate(period = "FY12")
fy13_wf <- extract_coeffs_demog(model2_13_wf) %>% mutate(period = "FY13")
fy14_wf <- extract_coeffs_demog(model2_14_wf) %>% mutate(period = "FY14")
fy15_wf <- extract_coeffs_demog(model2_15_wf) %>% mutate(period = "FY15")
fy16_wf <- extract_coeffs_demog(model2_16_wf) %>% mutate(period = "FY16")
fy17_wf <- extract_coeffs_demog(model2_17_wf) %>% mutate(period = "FY17")
fy18_wf <- extract_coeffs_demog(model2_18_wf) %>% mutate(period = "FY18")
fy19_wf <- extract_coeffs_demog(model2_19_wf) %>% mutate(period = "FY19")
fy20_wf <- extract_coeffs_demog(model2_20_wf) %>% mutate(period = "FY20")
fy21_wf <- extract_coeffs_demog(model2_21_wf) %>% mutate(period = "FY21")
fy22_wf <- extract_coeffs_demog(model2_22_wf) %>% mutate(period = "FY22")
fy23_wf <- extract_coeffs_demog(model2_23_wf) %>% mutate(period = "FY23")


bar_data_wf <- bind_rows(fy02_wf,
                          fy03_wf, 
                          fy04_wf,
                          fy05_wf,
                          fy06_wf,
                          fy07_wf,
                          fy08_wf,
                          fy09_wf,
                          fy10_wf,
                          fy11_wf,
                          fy12_wf,
                          fy13_wf, 
                          fy14_wf,
                          fy15_wf,
                          fy16_wf,
                          fy17_wf,
                          fy18_wf,
                          fy19_wf,
                          fy20_wf,
                          fy21_wf,
                          fy22_wf,
                          fy23_wf) 


bar_data_wf %>% filter(str_detect(demog, "Female")) %>% demogbar() 
  
```
  
  
```{r}
bar_data_wf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
<br>
<br>

### Hispanic Female Baseline

```{r}
model2_hf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype +
               whitecoll +
               upward +
               down +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 3) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}

model2_02_hf <- model2_hf(fy02)
model2_03_hf <- model2_hf(fy03)
model2_04_hf <- model2_hf(fy04)
model2_05_hf <- model2_hf(fy05)
model2_06_hf <- model2_hf(fy06)
model2_07_hf <- model2_hf(fy07)
model2_08_hf <- model2_hf(fy08)
model2_09_hf <- model2_hf(fy09)
model2_10_hf <- model2_hf(fy10)
model2_11_hf <- model2_hf(fy11)
model2_12_hf <- model2_hf(fy12)
model2_13_hf <- model2_hf(fy13)
model2_14_hf <- model2_hf(fy14)
model2_15_hf <- model2_hf(fy15)
model2_16_hf <- model2_hf(fy16)
model2_17_hf <- model2_hf(fy17)
model2_18_hf <- model2_hf(fy18)
model2_19_hf <- model2_hf(fy19)
model2_20_hf <- model2_hf(fy20)
model2_21_hf <- model2_hf(fy21)
model2_22_hf <- model2_hf(fy22)
model2_23_hf <- model2_hf(fy23)


```

```{r}

fy02_hf <- extract_coeffs_demog(model2_02_hf) %>% mutate(period = "FY02")
fy03_hf <- extract_coeffs_demog(model2_03_hf) %>% mutate(period = "FY03")
fy04_hf <- extract_coeffs_demog(model2_04_hf) %>% mutate(period = "FY04")
fy05_hf <- extract_coeffs_demog(model2_05_hf) %>% mutate(period = "FY05")
fy06_hf <- extract_coeffs_demog(model2_06_hf) %>% mutate(period = "FY06")
fy07_hf <- extract_coeffs_demog(model2_07_hf) %>% mutate(period = "FY07")
fy08_hf <- extract_coeffs_demog(model2_08_hf) %>% mutate(period = "FY08")
fy09_hf <- extract_coeffs_demog(model2_09_hf) %>% mutate(period = "FY09")
fy10_hf <- extract_coeffs_demog(model2_10_hf) %>% mutate(period = "FY10")
fy11_hf <- extract_coeffs_demog(model2_11_hf) %>% mutate(period = "FY11")
fy12_hf <- extract_coeffs_demog(model2_12_hf) %>% mutate(period = "FY12")
fy13_hf <- extract_coeffs_demog(model2_13_hf) %>% mutate(period = "FY13")
fy14_hf <- extract_coeffs_demog(model2_14_hf) %>% mutate(period = "FY14")
fy15_hf <- extract_coeffs_demog(model2_15_hf) %>% mutate(period = "FY15")
fy16_hf <- extract_coeffs_demog(model2_16_hf) %>% mutate(period = "FY16")
fy17_hf <- extract_coeffs_demog(model2_17_hf) %>% mutate(period = "FY17")
fy18_hf <- extract_coeffs_demog(model2_18_hf) %>% mutate(period = "FY18")
fy19_hf <- extract_coeffs_demog(model2_19_hf) %>% mutate(period = "FY19")
fy20_hf <- extract_coeffs_demog(model2_20_hf) %>% mutate(period = "FY20")
fy21_hf <- extract_coeffs_demog(model2_21_hf) %>% mutate(period = "FY21")
fy22_hf <- extract_coeffs_demog(model2_22_hf) %>% mutate(period = "FY22")
fy23_hf <- extract_coeffs_demog(model2_23_hf) %>% mutate(period = "FY23")


bar_data_hf <- bind_rows(fy02_hf,
                          fy03_hf, 
                          fy04_hf,
                          fy05_hf,
                          fy06_hf,
                          fy07_hf,
                          fy08_hf,
                          fy09_hf,
                          fy10_hf,
                          fy11_hf,
                          fy12_hf,
                          fy13_hf, 
                          fy14_hf,
                          fy15_hf,
                          fy16_hf,
                          fy17_hf,
                          fy18_hf,
                          fy19_hf,
                          fy20_hf,
                          fy21_hf,
                          fy22_hf,
                          fy23_hf) 


bar_data_hf %>% filter(str_detect(demog, "Female")) %>% demogbar() 
  
```
  
  
```{r}
bar_data_hf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```
<br>
<br>


### Black Female Baseline


```{r}
model2_bf <- function(df){
  lm(logsplit ~ logmin + 
               drugtraff +
               sexual2 + 
               porn +
               immigration +
               othtype +
               whitecoll +
               upward +
               down +
               subasst +
               mandmin2 +
               NEWCNVTN +
               custody +
               relevel(factor(racesex), ref = 1) +
               agedummy +
               educ +
               citizen,
             df)
}
```

```{r}

model2_02_bf <- model2_bf(fy02)
model2_03_bf <- model2_bf(fy03)
model2_04_bf <- model2_bf(fy04)
model2_05_bf <- model2_bf(fy05)
model2_06_bf <- model2_bf(fy06)
model2_07_bf <- model2_bf(fy07)
model2_08_bf <- model2_bf(fy08)
model2_09_bf <- model2_bf(fy09)
model2_10_bf <- model2_bf(fy10)
model2_11_bf <- model2_bf(fy11)
model2_12_bf <- model2_bf(fy12)
model2_13_bf <- model2_bf(fy13)
model2_14_bf <- model2_bf(fy14)
model2_15_bf <- model2_bf(fy15)
model2_16_bf <- model2_bf(fy16)
model2_17_bf <- model2_bf(fy17)
model2_18_bf <- model2_bf(fy18)
model2_19_bf <- model2_bf(fy19)
model2_20_bf <- model2_bf(fy20)
model2_21_bf <- model2_bf(fy21)
model2_22_bf <- model2_bf(fy22)
model2_23_bf <- model2_bf(fy23)

```

```{r}

fy02_bf <- extract_coeffs_demog(model2_02_bf) %>% mutate(period = "FY02")
fy03_bf <- extract_coeffs_demog(model2_03_bf) %>% mutate(period = "FY03")
fy04_bf <- extract_coeffs_demog(model2_04_bf) %>% mutate(period = "FY04")
fy05_bf <- extract_coeffs_demog(model2_05_bf) %>% mutate(period = "FY05")
fy06_bf <- extract_coeffs_demog(model2_06_bf) %>% mutate(period = "FY06")
fy07_bf <- extract_coeffs_demog(model2_07_bf) %>% mutate(period = "FY07")
fy08_bf <- extract_coeffs_demog(model2_08_bf) %>% mutate(period = "FY08")
fy09_bf <- extract_coeffs_demog(model2_09_bf) %>% mutate(period = "FY09")
fy10_bf <- extract_coeffs_demog(model2_10_bf) %>% mutate(period = "FY10")
fy11_bf <- extract_coeffs_demog(model2_11_bf) %>% mutate(period = "FY11")
fy12_bf <- extract_coeffs_demog(model2_12_bf) %>% mutate(period = "FY12")
fy13_bf <- extract_coeffs_demog(model2_13_bf) %>% mutate(period = "FY13")
fy14_bf <- extract_coeffs_demog(model2_14_bf) %>% mutate(period = "FY14")
fy15_bf <- extract_coeffs_demog(model2_15_bf) %>% mutate(period = "FY15")
fy16_bf <- extract_coeffs_demog(model2_16_bf) %>% mutate(period = "FY16")
fy17_bf <- extract_coeffs_demog(model2_17_bf) %>% mutate(period = "FY17")
fy18_bf <- extract_coeffs_demog(model2_18_bf) %>% mutate(period = "FY18")
fy19_bf <- extract_coeffs_demog(model2_19_bf) %>% mutate(period = "FY19")
fy20_bf <- extract_coeffs_demog(model2_20_bf) %>% mutate(period = "FY20")
fy21_bf <- extract_coeffs_demog(model2_21_bf) %>% mutate(period = "FY21")
fy22_bf <- extract_coeffs_demog(model2_22_bf) %>% mutate(period = "FY22")
fy23_bf <- extract_coeffs_demog(model2_23_bf) %>% mutate(period = "FY23")


bar_data_bf <- bind_rows(fy02_bf,
                          fy03_bf, 
                          fy04_bf,
                          fy05_bf,
                          fy06_bf,
                          fy07_bf,
                          fy08_bf,
                          fy09_bf,
                          fy10_bf,
                          fy11_bf,
                          fy12_bf,
                          fy13_bf, 
                          fy14_bf,
                          fy15_bf,
                          fy16_bf,
                          fy17_bf,
                          fy18_bf,
                          fy19_bf,
                          fy20_bf,
                          fy21_bf,
                          fy22_bf,
                          fy23_bf) 


bar_data_bf %>% filter(str_detect(demog, "Female")) %>% demogbar() 
  
``` 
  
```{r}
bar_data_bf %>% filter(str_detect(demog, "Male")) %>% demogbar()
```






```{r}
# New Data Vis


# bar_data_wf %>% filter(!str_detect(demog, "Other")) %>%  demogbar()
# bar_data_bf %>% filter(!str_detect(demog, "Other") & demog!="White Female") %>% demogbar()
# bar_data_hf %>% filter(!str_detect(demog, "Other") & str_detect(demog, "Female")) %>% demogbar()


```




